### 2014-12-03 manuscript/manuscript.Rnw
First commit. One figure added and part of the abstract.

\documentclass[a4paper]{article}

%--- Load packages. 
\usepackage{amssymb,amsmath}
\usepackage{framed}

% avoid ugly indentation of paragraphs
\usepackage{parskip}
\usepackage{graphicx}

% open sans font
\usepackage[default,osfigures,scale=0.95]{opensans}

% subfolder with the figures (within manuscript/)
\graphicspath{ {./figures/} }

% for author affiliations
\usepackage{authblk}

% for degree symbol
\usepackage{gensymb}

% for line numbers
\usepackage{lineno}

% margin size
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}

% for bibtex
\bibliographystyle{plain}

\begin{document}

% Turn line numbers on here.
%\linenumbers

\title{Catchy article title}

\author[1]{Remko Duursma}

\affil[1]{Hawkesbury Institute for the Environment}

\renewcommand\Authands{ and }
\maketitle

%--------------------------------------------------------------------------------------------%

<<preload, echo=FALSE>>=
library(knitr)
opts_chunk$set(echo=FALSE)
@

%--------------------------------------------------------------------------------------------%

\section*{Summary}

1. The \texttt{plantecophys} R package includes many functions for analysis and modelling of plant leaf gas exchange data.

2. This paper describes the new package, and gives examples on its use in plant ecophysiological research. The package makes technically difficult calculations easily accessible for many users.

3. We briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of A-Ci curves, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.

\section*{Introduction}

Time to cite some papers, some folks said so \cite{schenk_rooting_2002}, and some others as well \cite{mokany_critical_2006}.

%--------------------------------------------------------------------------------------------%

\clearpage
\section{Figures}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{figure1.pdf}
    \caption{Standard output from the \texttt{fitaci} function. $A_net$ is the net photosynthetic rate, $C_i$ the intercellular CO$_2$ concentration. Symbols are measurements, the line the fitted Farquhar model of photosynthesis.}
    \label{fig:figure1}
\end{figure}

\clearpage
\bibliography{references}

\end{document}

### 2014-12-03 R/functions.R

@@ -1,0 +1,18 @@
+to.pdf <- function(expr, filename, ..., verbose=TRUE) {
+  if(!file.exists(dirname(filename)))
+    dir.create(dirname(filename), recursive=TRUE)
+  if ( verbose )
+    cat(sprintf("Creating %s\n", filename))
+  pdf(filename, ...)
+  on.exit(dev.off())
+  eval.parent(substitute(expr))
+}
+to.eps <- function(expr, filename, ..., verbose=TRUE) {
+  if(!file.exists(dirname(filename)))
+    dir.create(dirname(filename), recursive=TRUE)
+  if ( verbose )
+    cat(sprintf("Creating %s\n", filename))
+  postscript(filename, ...)
+  on.exit(dev.off())
+  eval.parent(substitute(expr))
+}

### 2014-12-03 figures.R

@@ -1,0 +1,14 @@
+# Figures
+source("R/functions.R")
+library(plantecophys)
+
+figure1 <- function(){
+  
+  
+  f <- fitaci(acidata1)
+  
+  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1)
+  plot(f)
+  
+}
+to.pdf(figure1(), filename="manuscript/figures/figure1.pdf")

### 2014-12-03 manuscript/manuscript.Rnw
Added references and some text

@@
 % for bibtex
-\bibliographystyle{plain}
+\usepackage{authordate1-4}
+\bibliographystyle{authordate1}
 
@@
 
-\title{Catchy article title}
+\title{\textbf{plantecophys} : an R package for analysis and modelling of leaf gas exchange data}
 
@@
 
-Time to cite some papers, some folks said so \cite{schenk_rooting_2002}, and some others as well \cite{mokany_critical_2006}.
+The Farquhar model of photosynthesis is widely used \cite{Farquhar1980}.
+Various methods exist to fit the model to data.
+We have implemented \texttt{fitaci}, which fits the model in two different ways.
+The function \texttt{Aci} simulates the A-Ci curve.
+
+The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO$_2$ concentration 
+\cite{Farquhar1982}. This is solved by the \texttt{Photosyn} function.
+
+Implementation of several stomatal conductance models, including \cite{ballberry1987,leuning1995,medlyn2011}.
+
+We have used the code in several of our papers, including \cite{medlyn2011,duursma2013,duursma2014}
+
+Full numerical optimizatin of the Cowan-Farquhar hypothesis. We have used this code in \cite{medlyn2011}, and is very similar to that used by \cite{buckley2014}.
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{figure2.pdf}
+    \caption{The intersection of the supply and demand curves of photosynthesis as solved by \texttt{Photosyn}.}
+    \label{fig:figure1}
+\end{figure}
+
 \clearpage

### 2014-12-03 figures.R

@@ -12,3 +12,35 @@ figure1 <- function(){
   
 }
 to.pdf(figure1(), filename="manuscript/figures/figure1.pdf")
+
+
+
+# Sharkey figure
+figure2 <- function(){
+  ci <- seq(50, 800, length=101)
+  
+  acirun <- Aci(Ci=ci)
+  
+  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
+  with(acirun, plot(Ci, ALEAF, type='l',
+                    ylim=c(0,20), xlim=c(0,800),
+                    col="blue",lwd=2,
+                    xlab=expression(italic(C)[i]~~(ppm)),
+                    ylab=expression(italic(A)[leaf]~~(mu*mol~m^-2~s^-1))))
+  
+  p <- Photosyn(VPD=2.5, g1=2)
+  
+  gc <- p$GS / 1.6
+  abline(gc*p$Ca, -gc, col="forestgreen", lwd=2)
+  
+  points(p$Ci, p$ALEAF, pch=19, col="red", cex=1.1)
+  
+  legend("right", c(expression("Demand curve: "~A==f(italic(C)[i])),
+                          "Supply curve: "~A==g[CO2]*(italic(C)[a]-italic(C)[i]),
+                          "Operating point"),
+         lwd=c(2,2,-1),pch=c(-1,-1,19), pt.cex=c(-1,-1,1.1),
+         col=c("blue","forestgreen","red"), inset=0.01, cex=1.1, bty='n')
+}
+to.pdf(figure2(), filename="manuscript/figures/figure2.pdf")
+
+  

### 2015-05-22 manuscript.Rmd
made switch to markdown

---
title: plantecophys - an R package for analysis and modelling of leaf gas exchange data
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: plos-one.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
*Corresponding author*: 
Remko Duursma
E: remkoduursma@gmail.com
T: +61(0)422096908

```{r echo=FALSE, message=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# mine
library(plantecophys)

# figures
source("figures.R")

# Do calculations here, create objects that we can use throughout the paper.

# Or, place them all in a separate script:
# source("manuscript_calcs.R")

# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))

# Nice function to format P values.
pval <- function(...)formatPval(..., includeEquality=TRUE)

# Set Flag to include figures in text or not.
includeFigs <- TRUE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")
```

# Abstract {.unnumbered}

1. The plantecophys R package includes many functions for analysis and modelling of plant leaf gas exchange data.

2. This paper describes the new package, and gives examples on its use in plant ecophysiological research. The package makes technically difficult calculations easily accessible for many users.

3. We briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of A-Ci curves, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.

# Introduction

The Farquhar model of photosynthesis is widely used [@Farquhar1980]. Various methods exist to fit the model to data. We have implemented `fitaci`, which fits the model in two different ways. The function `Aci` simulates the A-Ci curve.

The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration [@Farquhar1982]. This is solved by the `Photosyn` function.

Implementation of several stomatal conductance models, including [@ballberry1987; @leuning1995; @medlyn2011].

We have used the code in several of our papers, including [@medlyn2011; @duursma2013; @duursma2014].

Full numerical optimization of the Cowan-Farquhar hypothesis. We have used this code in [@medlyn2011], and is very similar to that used by [@buckley2014].

# Design and implementation

## Subheader text
Curabitur suscipit, orci ac luctus pellentesque, lacus quam lobortis leo, sit amet rhoncus arcu dolor in felis. Morbi suscipit ullamcorper velit, vitae facilisis ligula tincidunt eu. Mauris molestie metus metus, sit amet eleifend mi volutpat sit amet. Nam dignissim ipsum eu tellus malesuada dignissim. Integer eleifend id metus vitae eleifend. Aliquam ultricies mi a eros vulputate, sit amet venenatis ante suscipit. Morbi vehicula sed nisi vitae semper. In et libero ac magna tincidunt mattis. Duis et enim eu dui scelerisque ullamcorper. Praesent pulvinar erat vitae bibendum iaculis. Maecenas dignissim ante nec dui vehicula dictum. Donec ac sodales arcu. Morbi nec lacus luctus, ultrices lacus vitae, rutrum justo.

## More methods subheader
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim,

# Results

Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim.

# Availability and future directions

Curabitur semper nisl eu turpis tempus molestie. Suspendisse potenti. Proin laoreet nibh ac nisi iaculis, at efficitur augue tincidunt. Cras mi lectus, maximus a venenatis eget, pellentesque vel erat. Cras tincidunt, quam ultricies feugiat condimentum, velit ligula tincidunt velit, id venenatis lacus risus id sapien. Ut pulvinar quam ut faucibus ultricies. Ut quis enim et eros hendrerit dictum. Nunc tempus magna vel ante convallis pretium. Praesent sit amet risus vestibulum justo ultrices ultricies vel eu metus. In aliquam ultrices diam, ut tempus mi ultrices ut. Sed id lacus ac sem ullamcorper maximus ut eu leo.

\
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim.

# Acknowledgements {.unnumbered}
We thank lots of people.

# Figures {.unnumbered}

```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
figure1()
```

**Figure 1**. Standard output from the `fitaci` function. A~net~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration. Symbols are measurements, the line the fitted Farquhar model of photosynthesis.

```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
figure2()
```

** Figure 2**. The intersection of the supply and demand curves of photosynthesis as solved by `Photosyn`.

# References {.unnumbered}

### 2015-05-22 figures.R

@@ -11,7 +11,6 @@ figure1 <- function(){
   plot(f)
   
 }
-to.pdf(figure1(), filename="manuscript/figures/figure1.pdf")
 
 
 
@@ -41,6 +40,5 @@ figure2 <- function(){
          lwd=c(2,2,-1),pch=c(-1,-1,19), pt.cex=c(-1,-1,1.1),
          col=c("blue","forestgreen","red"), inset=0.01, cex=1.1, bty='n')
 }
-to.pdf(figure2(), filename="manuscript/figures/figure2.pdf")
 
   

### 2015-05-22 manuscript.Rmd
cleanup after switch to markdown

@@
 T: +61(0)422096908
+
+<!---
+For Submission to PLOS one
+e.g. http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0061217
+http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003929
+http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003270
+and see: http://journals.plos.org/plosone/s/submission-guidelines
+(Methods, software, databases, and tools)
+--->
 

### 2015-06-24 manuscript.Rmd
started writing the thing

@@
 
-1. The plantecophys R package includes many functions for analysis and modelling of plant leaf gas exchange data.
+1. The plantecophys R package includes many functions for analysis and modelling of plant leaf gas exchange data, as commonly collected with a portable gas exchange instrument.
 
-2. This paper describes the new package, and gives examples on its use in plant ecophysiological research. The package makes technically difficult calculations easily accessible for many users.
+2.
+This paper describes the functionality and implementation of the new package, and gives examples on its use in plant ecophysiological research.
+The package makes technically difficult calculations easily accessible for many users.
 
 3.
-We briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of A-Ci curves, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
+I briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
 
@@
 
+Wealth of leaf gas exchange data collected in the field of plant ecological physiology. Especially since the advent of portable, easy to use, gas exchange instruments such as the Licor-6400.
+
+These data are used to estimate important plant traits such as Vcmax&Jmax, as well to study responses of stomatal conductance to key environmental drivers.
+
 The Farquhar model of photosynthesis is widely used [@Farquhar1980].
-Various methods exist to fit the model to data.
-We have implemented `fitaci`, which fits the model in two different ways.
-The function `Aci` simulates the A-Ci curve.
+Various methods exist to fit the model to data, which requires some finesse because the model is not a single non-linear equation, but instead a minimum function that is sometimes difficult to fit (Gu), and sample sizes collected are often small due to time constraints in the field.
+We do not repeat a description of the Farquhar model here, it has been described many times.
+But generally it is of the form,
 
-The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration [@Farquhar1982]. This is solved by the `Photosyn` function.
+(1)
+$$A_n = min(A_c, A_j) - R_d$$
+
+where A~c~ is the gross photosynthesis rate when Rubisco activity is limiting, A~j~ when RuBP-regeneration is limiting, and R~d~ the rate of dark respiration.
+A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i\Gamma^∗)/(k_2 + C_i)$, where $\Gamma^∗$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
+The details of these functions and the temperature dependence of the parameters are described elsewhere (e.g.
+Medlyn et al., 2002).
+As mentioned above, the parameters Vcmax and Jmax that are embedded in Eq.
+1 are not easy to estimate, moreover because the various dependencies of the parameters on leaf temperature must be accounted for.
+
+Once the A-Ci curve is fitted, we can use the Farquhar (FvC) model to simulate the relationship between substrate availability (for which Ci is as a proxy) and rates of photosynthetic CO2 assimilation.
+The intercellular CO2 concentration, in turn, is set by stomatal opening, which modulates the diffusion of CO2 into leaves via Fick's law:
+
+(2)
+$$ A_n = g_s(C_a - C_i) $$
+
+We now have two equations for the photosynthesis rate.
+The demand-based one (Eq.
+1) and the diffusion-based one.
+At steady state these two should be equal, which can be graphically shown as in Fig.
+1 (cf.
+[@Farquhar1982]).
+
+The next step is to include a model of g~s~ that allows for the known response of g~s~ to drivers including vapour pressure deficit and atmospheric CO~2~ concentration (C~a~).
+The most successful, though empirical, suite of gs models is the Ball-Berry class of models where gs is modelled as a function of A~n~ and D and CO2.
+This way, effects of leaf temperature and PPFD - both of which are known to affect gs - are modelled through the dependency of photosynthesis on these drivers.
+Thus the third equation is,
+
+(3)
+$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$
+
+Through Eqs.
+1-3 we now have the coupled leaf gas exchange model, which allows modelling of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g1, Vcmax, Jmax, Rd).
+This coupled leaf gas exchange model can be solved as a quadratic equation, yielding the setpoint C~i~ (Fig.
+1), and has found use in many applications (cite).
+In the next section we describe the various options for f(D) as well as an independent approach assuming that stomatal conductance is optimized to maximize photosynthesis at a given water expenditure (cf.
+Cowan, Medlyn).
+
+Despite the widespread use of a standard instrument to collect data, tools to analyze the data are scattered and subject to little standardization.
+Some excel sheets here and there, etc., but we need a solid set of tools in the widely used R language.
+I here describe the plantecophys package, implemented in the R language.
+
+# Design and implementation
+
+## The main functions
+We have implemented `fitaci`, which fits the model in two different ways. The function `Aci` simulates the A-Ci curve.
+
+The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration. This is solved by the `Photosyn` function.
 
@@
 Full numerical optimization of the Cowan-Farquhar hypothesis. We have used this code in [@medlyn2011], and is very similar to that used by [@buckley2014].
-
-# Design and implementation
-
-## Subheader text
-Curabitur suscipit, orci ac luctus pellentesque, lacus quam lobortis leo, sit amet rhoncus arcu dolor in felis.
-Morbi suscipit ullamcorper velit, vitae facilisis ligula tincidunt eu.
-Mauris molestie metus metus, sit amet eleifend mi volutpat sit amet.
-Nam dignissim ipsum eu tellus malesuada dignissim.
-Integer eleifend id metus vitae eleifend.
-Aliquam ultricies mi a eros vulputate, sit amet venenatis ante suscipit.
-Morbi vehicula sed nisi vitae semper.
-In et libero ac magna tincidunt mattis.
-Duis et enim eu dui scelerisque ullamcorper.
-Praesent pulvinar erat vitae bibendum iaculis.
-Maecenas dignissim ante nec dui vehicula dictum.
-Donec ac sodales arcu.
-Morbi nec lacus luctus, ultrices lacus vitae, rutrum justo.
 
@@
 ```
+** Figure 1**.
+The intersection of the supply and demand curves of photosynthesis.
+The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvC model) are known.
 
-**Figure 1**.
+```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+figure2()
+```
+**Figure 2**.
 Standard output from the `fitaci` function.
@@
 
-```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure2()
-```
-
-** Figure 2**. The intersection of the supply and demand curves of photosynthesis as solved by `Photosyn`.
-
 # References {.unnumbered}

### 2015-06-24 figures.R

@@ -2,20 +2,10 @@
 source("R/functions.R")
 library(plantecophys)
 
-figure1 <- function(){
-  
-  
-  f <- fitaci(acidata1)
-  
-  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1)
-  plot(f)
-  
-}
-
 
 
 # Sharkey figure
-figure2 <- function(){
+figure1 <- function(){
   ci <- seq(50, 800, length=101)
   
   acirun <- Aci(Ci=ci)
@@ -39,6 +29,18 @@ figure2 <- function(){
                           "Operating point"),
          lwd=c(2,2,-1),pch=c(-1,-1,19), pt.cex=c(-1,-1,1.1),
          col=c("blue","forestgreen","red"), inset=0.01, cex=1.1, bty='n')
+}
+
+
+# example output fitaci
+figure2 <- function(){
+  
+  
+  f <- fitaci(acidata1)
+  
+  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1)
+  plot(f)
+  
 }
 
   

### 2015-06-25 manuscript.Rmd
progress

@@
 
-The Farquhar model of photosynthesis is widely used [@Farquhar1980].
+The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is very widely used in interpreting leaf gas exchange data (cites), up to predicting carbon uptake by vegetation at a global scale.
+The key prediction of the model is the response of photosynthesis to [CO~2~] in the intercellular airspaces (Ci).
+In turn, Ci is controlled the result of the balance between diffusion of CO2 through the stomata, and the uptake of CO2 by photosynthesis.
+To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO2] concentrations.
+
 Various methods exist to fit the model to data, which requires some finesse because the model is not a single non-linear equation, but instead a minimum function that is sometimes difficult to fit (Gu), and sample sizes collected are often small due to time constraints in the field.
@@
 Once the A-Ci curve is fitted, we can use the Farquhar (FvC) model to simulate the relationship between substrate availability (for which Ci is as a proxy) and rates of photosynthetic CO2 assimilation.
-The intercellular CO2 concentration, in turn, is set by stomatal opening, which modulates the diffusion of CO2 into leaves via Fick's law:
 
@@
 1 (cf.
-[@Farquhar1982]).
+[@farquhar1982]).
 
@@
 
+We have used the code in several of our papers, including [@medlyn2011; @duursma2013; @duursma2014].
+
 # Design and implementation
@@
 ## The main functions
-We have implemented `fitaci`, which fits the model in two different ways. The function `Aci` simulates the A-Ci curve.
+
+We have implemented `fitaci`, which fits the FvC model in two different ways.
+A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that both prevents batch analysis and is arbitrary.
+Another method (http://leafweb.ornl.gov/) requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
+
+Details about it : fits the hyperbolic min function simultaneously.
+Retrieves estimates of Vcmax, Jmax and Rd (the latter can also be held constant at a known value, in which case only Jmax and Vcmax are fit).
+Optionally can specify the transition point, so that the Vcmax-limited part is fit independent of the Jmax-limited part.
+Returns parameter estimates and standard errors, and includes a plot method.
+Fig 2.
+is made like this:
+
+```
+f <- fitaci(mydata)
+plot(f)
+```
+
+The function `Aci` simulates the A-Ci curve.
 
@@
 
-We have used the code in several of our papers, including [@medlyn2011; @duursma2013; @duursma2014].
-
 Full numerical optimization of the Cowan-Farquhar hypothesis. We have used this code in [@medlyn2011], and is very similar to that used by [@buckley2014].
 
-## More methods subheader
-Aliquam ut ligula at nulla mattis rutrum.
-Duis dictum tempus diam ac tincidunt.
-Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero.
-In convallis sem id dui blandit, id placerat libero molestie.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae.
-Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt.
-Morbi ut ligula vitae felis convallis cursus.
-Fusce placerat nisl et mi imperdiet viverra.
-In semper nibh sollicitudin tortor commodo pretium.
-Vestibulum ultricies dignissim sollicitudin.
-Vivamus tristique pellentesque ante vitae tempor.
-Morbi facilisis urna vitae sapien blandit feugiat.
-Integer dapibus ut ex id dignissim,
+## Example analysis
 
-# Results
-
-Aliquam ut ligula at nulla mattis rutrum.
-Duis dictum tempus diam ac tincidunt.
-Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero.
-In convallis sem id dui blandit, id placerat libero molestie.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae.
-Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt.
-Morbi ut ligula vitae felis convallis cursus.
-Fusce placerat nisl et mi imperdiet viverra.
-In semper nibh sollicitudin tortor commodo pretium.
-Vestibulum ultricies dignissim sollicitudin.
-Vivamus tristique pellentesque ante vitae tempor.
-Morbi facilisis urna vitae sapien blandit feugiat.
-Integer dapibus ut ex id dignissim.
+fit aci curve, estimate g1, model photosynthesis and transpiration for a met dataset?
 
@@
 
-Curabitur semper nisl eu turpis tempus molestie.
-Suspendisse potenti.
-Proin laoreet nibh ac nisi iaculis, at efficitur augue tincidunt.
-Cras mi lectus, maximus a venenatis eget, pellentesque vel erat.
-Cras tincidunt, quam ultricies feugiat condimentum, velit ligula tincidunt velit, id venenatis lacus risus id sapien.
-Ut pulvinar quam ut faucibus ultricies.
-Ut quis enim et eros hendrerit dictum.
-Nunc tempus magna vel ante convallis pretium.
-Praesent sit amet risus vestibulum justo ultrices ultricies vel eu metus.
-In aliquam ultrices diam, ut tempus mi ultrices ut.
-Sed id lacus ac sem ullamcorper maximus ut eu leo.
-
-\
-Aliquam ut ligula at nulla mattis rutrum.
-Duis dictum tempus diam ac tincidunt.
-Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero.
-In convallis sem id dui blandit, id placerat libero molestie.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae.
-Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt.
-Morbi ut ligula vitae felis convallis cursus.
-Fusce placerat nisl et mi imperdiet viverra.
-In semper nibh sollicitudin tortor commodo pretium.
-Vestibulum ultricies dignissim sollicitudin.
-Vivamus tristique pellentesque ante vitae tempor.
-Morbi facilisis urna vitae sapien blandit feugiat.
-Integer dapibus ut ex id dignissim.
+The package is freely available under the XXX license (check CRAN). The development version is under version control at bitbucket. New changes are made available frequently.
 
 # Acknowledgements {.unnumbered}
-We thank lots of people.
+Thanks to John Drake for extensive testing, David Ellsworth, Belinda Medlyn, Martin de Kauwe for discussion on fitting A-Ci curves.
 

### 2015-07-02 manuscript.Rmd
more writing

@@
 
-1. The plantecophys R package includes many functions for analysis and modelling of plant leaf gas exchange data, as commonly collected with a portable gas exchange instrument.
-
-2.
-This paper describes the functionality and implementation of the new package, and gives examples on its use in plant ecophysiological research.
-The package makes technically difficult calculations easily accessible for many users.
-
-3.
+Here I present the R package 'plantecophys', a toolkit to analyse and model leaf gas exchange data, a cornerstone of plant ecophysiological research.
+Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models.
+These models include the Farquhar-von Caemmerer-Berry model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO2 in the leaf.
+The 'plantecophys' R package includes functions both for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data.
+Here I describe the functionality and implementation of the new package, and give examples on its use in plant ecophysiological research.
+The package makes technically challenging calculations easily accessible for many users.
 I briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
+The package is freely available on CRAN, and a development version is hosted at http://www.bitbucket.org/remkoduursma/plantecophys.
 
@@
 
+The main functions in the package include functions to :
+
+- Fit, summarize, plot and simulate photosynthesis-[CO~2~] response curves (known as 'A-Ci' curves)
+- Fit Ball-Berry type models of stomatal conductance
+- Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis
+- Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model
+- Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present
+- Convert between commonly used units (relative humidity, vapour pressure deficit, dewpoint temperature).
+
+## Language
+
+The 'plantecophys' package is implemented in native R, and has no dependencies on other packages.
+As such it compiles easily, and is highly portable.
+The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys).
+The repository includes an issue tracker, where users can suggest changes or report bugs.
+
+## Results
+
+### Fitting A-Ci curves
+
 We have implemented `fitaci`, which fits the FvC model in two different ways.
@@
 Another method (http://leafweb.ornl.gov/) requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
+Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see e.g.
+https://github.com/mdekauwe/FitFarquharModel).
 
@@
 
-The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration. This is solved by the `Photosyn` function.
+### Fitting stomatal conductance models
 
@@
 
+### Coupled leaf gas exchange model
+
+The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration. This is solved by the `Photosyn` function.
+
+### Numerical solution of optimal stomatal conductance
+
 Full numerical optimization of the Cowan-Farquhar hypothesis. We have used this code in [@medlyn2011], and is very similar to that used by [@buckley2014].
 
-## Example analysis
+# Conclusions
 
-fit aci curve, estimate g1, model photosynthesis and transpiration for a met dataset?
+We need a publicly available, open source, set of tools to analyze these data which form the cornerstone of plant physiological ecology.
+Though alternative methods exist to fit A-Ci curves and other details of the implementations can see alternatives, the plantecophys package provides a platform where such methods comparisons can be made.
+The R language is now widely used, and many users can make changes to the code and recompile the package.
+We have included a set of functions that are key to ecophysiological research. There is certainly scope, and interest, to increase the array of tools included in the package.
 
-# Availability and future directions
+# Availability and requirements
 
-The package is freely available under the XXX license (check CRAN). The development version is under version control at bitbucket. New changes are made available frequently.
+**Project name**: plantecophys
+
+**Project Stable Release**: cran.r-project.org/package=plantecophys
+
+**Project Home Page**: http://www.bitbucket.org/remkoduursma/plantecophys
+
+**Project Issue Tracker**: http://www.bitbucket.org/remkoduursma/plantecophys/issues
+
+**Operating System(s)**: Platform Independent
+
+**Programming Language(s)**: R
+
+**Other Requirements**: none
 
 # Acknowledgements {.unnumbered}
-Thanks to John Drake for extensive testing, David Ellsworth, Belinda Medlyn, Martin de Kauwe for discussion on fitting A-Ci curves.
+Thanks to John Drake for extensive testing.
+David Ellsworth, Belinda Medlyn, Martin de Kauwe are acknowledged for discussion on fitting A-Ci curves.
+Special thanks to Belinda Medlyn whose code in the MAESTRA model provided inspiration for an early version of the Photosyn function.
 

### 2015-07-06 manuscript.Rmd
New figures, lots of text updated.

@@
 ---
-title: plantecophys - an R package for analysis and modelling of leaf gas exchange data
+title: plantecophys - an R package for analysis and modelling leaf gas exchange data
 author: ''
@@
 Here I describe the functionality and implementation of the new package, and give examples on its use in plant ecophysiological research.
-The package makes technically challenging calculations easily accessible for many users.
 I briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
-The package is freely available on CRAN, and a development version is hosted at http://www.bitbucket.org/remkoduursma/plantecophys.
+The package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.
 
@@
 
-Wealth of leaf gas exchange data collected in the field of plant ecological physiology. Especially since the advent of portable, easy to use, gas exchange instruments such as the Licor-6400.
+Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO2 and H~2~O has been collected.
+These data play a central role in physiological plant ecology, because rates of photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
+Not only do leaf gas exchange data allow detailed studies of the underlying plant physiological controls, they are also used to parameterize an important component of land surface models used to predict global water and carbon cycling [e.g.
+@bonan2014; @dekauwe2015].
 
-These data are used to estimate important plant traits such as Vcmax&Jmax, as well to study responses of stomatal conductance to key environmental drivers.
+\
+The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting leaf gas exchange data (cites), by providing sensible and comparable metrics of the photosynthetic capacity, and accurately predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
+This widely cited model is embedded in many global vegetation models (cite).
+The key prediction of the model is the response of photosynthesis to [CO~2~] in the intercellular airspaces (Ci).
+To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO2] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~).
 
-The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is very widely used in interpreting leaf gas exchange data (cites), up to predicting carbon uptake by vegetation at a global scale.
-The key prediction of the model is the response of photosynthesis to [CO~2~] in the intercellular airspaces (Ci).
-In turn, Ci is controlled the result of the balance between diffusion of CO2 through the stomata, and the uptake of CO2 by photosynthesis.
-To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO2] concentrations.
-
-Various methods exist to fit the model to data, which requires some finesse because the model is not a single non-linear equation, but instead a minimum function that is sometimes difficult to fit (Gu), and sample sizes collected are often small due to time constraints in the field.
-We do not repeat a description of the Farquhar model here, it has been described many times.
-But generally it is of the form,
+\
+We do not repeat a detailed description of the FvC model here, as it has been described many times. But generally it is of the form,
 
@@
 
+\
 where A~c~ is the gross photosynthesis rate when Rubisco activity is limiting, A~j~ when RuBP-regeneration is limiting, and R~d~ the rate of dark respiration.
 A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i\Gamma^∗)/(k_2 + C_i)$, where $\Gamma^∗$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
-The details of these functions and the temperature dependence of the parameters are described elsewhere (e.g.
-Medlyn et al., 2002).
-As mentioned above, the parameters Vcmax and Jmax that are embedded in Eq.
-1 are not easy to estimate, moreover because the various dependencies of the parameters on leaf temperature must be accounted for.
+The details of these functions and the temperature dependence of the parameters are described elsewhere [@medlyn2002].
+The use of C~i~ in these equations is routine, but actually it should be Cc, considering that CO2 in the chloroplast is not equal to that in the intercellular air space.
 
-Once the A-Ci curve is fitted, we can use the Farquhar (FvC) model to simulate the relationship between substrate availability (for which Ci is as a proxy) and rates of photosynthetic CO2 assimilation.
+\
+Fitting the model to data requires some finesse because the model is not a single non-linear equation, but instead a minimum function of two non-linear equations that is sometimes difficult to fit.
+Moreover, sample sizes collected are often small due to time constraints in the field.
+A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that both prevents batch analysis and is arbitrary.
+Another method (http://leafweb.ornl.gov/) requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
+Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]).
+
+\
+Through Eq.
+1, we have a dependency of photosynthesis on the availability of the substrate, Ci.
+To estimate Ci itself, we must have an estimate of the stomatal conductance to CO~2~.
+From Fick's law, we can relate A~n~ to g~s~ and C~i~ as,
 
@@
 
+\
+where g~s~ is the conductance to CO~2~.
 We now have two equations for the photosynthesis rate.
-The demand-based one (Eq.
-1) and the diffusion-based one.
-At steady state these two should be equal, which can be graphically shown as in Fig.
+The 'demand function' is given by Eq.
+1, and the 'supply function' in Eq.
+2.
+At steady state these two equation should be equal, which can be graphically shown as in Fig.
 1 (cf.
@@
 
-The next step is to include a model of g~s~ that allows for the known response of g~s~ to drivers including vapour pressure deficit and atmospheric CO~2~ concentration (C~a~).
-The most successful, though empirical, suite of gs models is the Ball-Berry class of models where gs is modelled as a function of A~n~ and D and CO2.
-This way, effects of leaf temperature and PPFD - both of which are known to affect gs - are modelled through the dependency of photosynthesis on these drivers.
+\
+Because g~s~ itself responds to environmental drivers, we must then estimate this dependence with another model.
+The most successful, though empirical, g~s~ models is the Ball-Berry [@ballberry1987] class of models.
+This model posits an entirely empirical equation that describes the response of g~s~ to air humidity, CO~2~ and the rate of photosynthesis.
+This way, effects of leaf temperature and PPFD - both of which are known to affect g~s~ - are modelled through the dependency of A~n~ on these drivers.
 Thus the third equation is,
@@
 
-Through Eqs.
-1-3 we now have the coupled leaf gas exchange model, which allows modelling of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g1, Vcmax, Jmax, Rd).
-This coupled leaf gas exchange model can be solved as a quadratic equation, yielding the setpoint C~i~ (Fig.
-1), and has found use in many applications (cite).
-In the next section we describe the various options for f(D) as well as an independent approach assuming that stomatal conductance is optimized to maximize photosynthesis at a given water expenditure (cf.
-Cowan, Medlyn).
+\
+where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters.
+An alternative approach to modelling g~s~ is through the hypothesis that stomata act optimally in the sense that they maximize photosynthesis while minimizing water loss.
+This hypothesis was first developed by Cowan and Farquhar [@cowan1977] and has seen many applications.
+Medlyn et al.
+[@medlyn2011] showed that the optimality hypothesis, when coupled to the FvC model, leads to an expression analogous to the Ball-Berry type models (Eq.
+3), but with a different D response function (f(D) in Eq.
+3) than the original Ball-Berry model, or yet another version introduced by Leuning [@leuning1995].
 
-Despite the widespread use of a standard instrument to collect data, tools to analyze the data are scattered and subject to little standardization.
-Some excel sheets here and there, etc., but we need a solid set of tools in the widely used R language.
-I here describe the plantecophys package, implemented in the R language.
+\
+Finally we can combine the biochemical demand function of photosynthesis (Eq.
+1) with the supply function (Eq.
+2) and an expression for the dependency of g~s~ on environmental drivers (Eq.
+3).
+This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991] is implemented in many process-based ecosystem models [@wangl1998; @bonan2014].
+This model allows prediction of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g1, Vcmax, Jmax, Rd).
 
-We have used the code in several of our papers, including [@medlyn2011; @duursma2013; @duursma2014].
+\
+Despite the widespread use of the FvC model and the coupled leaf gas exchange model, tools to analyze the data are scattered and subject to little standardization.
+Fitting the FvC model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
+The coupled leaf gas exchange model is not straightforward to implement, and I do not know of an open-source implementation that is publicly available.
+I here describe the plantecophys package, implemented in the R language [@rfoundation].
+The code is freely available (without restrictions), and managed with a version control system.
+I have developed the code base in my work on leaf and canopy modelling of photosynthesis and stomatal conductance, with many additions based on user requests [@medlyn2011; @barton2012; @medlyn2013; @duursma2013; @duursma2014].
 
@@
 
-The main functions in the package include functions to :
+The main functions in the package are summarized in Table 1.
 
-- Fit, summarize, plot and simulate photosynthesis-[CO~2~] response curves (known as 'A-Ci' curves)
-- Fit Ball-Berry type models of stomatal conductance
-- Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis
-- Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model
-- Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present
-- Convert between commonly used units (relative humidity, vapour pressure deficit, dewpoint temperature).
+```{r results='asis', echo=FALSE, message=FALSE, warn=FALSE}
+f <- c("`fitaci`","`fitBB`","`FARAO`","`Photosyn`","`PhotosynEB`","`RHtoVPD` etc.")
+d <- c("Fit, summarize, plot and simulate photosynthesis-[CO~2~] response curves (A-Ci curves)",
+"Fit Ball-Berry type models of stomatal conductance",
+"Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
+"Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model",
+"Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
+"Convert between commonly used units (relative humidity, vapour pressure deficit, dewpoint temperature)")
+df <- data.frame(Function=f, Description=d)
+
+# library(knitr)
+# kable(df, format="markdown",
+#       caption="Table 1. Main functions in the plantecophys package. Each function has its own detailed help page  including example code.")
+
+library(ascii)
+suppressWarnings(print(ascii(df, caption="Table 1. Main functions in the plantecophys package.",
+            include.rownames=FALSE),
+      type='pandoc'))
+```
 
@@
 The 'plantecophys' package is implemented in native R, and has no dependencies on other packages.
-As such it compiles easily, and is highly portable.
+As such it builds easily, and is highly portable.
 The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys).
@@
 
-## Results
+## Results and Discussion
 
@@
 
-We have implemented `fitaci`, which fits the FvC model in two different ways.
-A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that both prevents batch analysis and is arbitrary.
-Another method (http://leafweb.ornl.gov/) requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
-Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see e.g.
-https://github.com/mdekauwe/FitFarquharModel).
+The `fitaci` function fits the FvC model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors.
+Optionally R~d~ can be provided as a known value, otherwise it is estimated from the A-Ci curve.
+The user does not have to provide the transition point (see Fig.
+1), as this is estimated by `fitaci` automatically.
+It is however an option to fix the transition point, which may be helpful to check whether the best fit was achieved.
+Standard errors are estimated with the usual methods for non-linear regression (`nls` function in base R).
+Instead of fitting the minimum function (Eq.
+1) it is actually far more practical to use a hyperbolic minimum, which avoids a discontinuity at the transition from A~c~ to A~j~ limitation.
 
-Details about it : fits the hyperbolic min function simultaneously.
-Retrieves estimates of Vcmax, Jmax and Rd (the latter can also be held constant at a known value, in which case only Jmax and Vcmax are fit).
-Optionally can specify the transition point, so that the Vcmax-limited part is fit independent of the Jmax-limited part.
-Returns parameter estimates and standard errors, and includes a plot method.
-Fig 2.
-is made like this:
+\
+The function takes a dataframe as input, which includes A~n~, C~i~ and optionally leaf temperature (T~leaf~) and photosynthetically active radiation (PAR).
+It then returns a special object which has a print and plot method.
 
 ```
+# Fit FvC model
 f <- fitaci(mydata)
+
+# Make standard plot
 plot(f)
@@
 
-The function `Aci` simulates the A-Ci curve.
+\
+The output of the above example is shown in Fig.
+2.
+Additionally, `fitacis` can be sued to fit many curves at once, for example one for each species or site in a dataset.
+The function `Aci` simulates the A-Ci curve, that is, it returns A~n~ when C~i~ is known.
 
@@
 
-Implementation of several stomatal conductance models, including [@ballberry1987; @leuning1995; @medlyn2011].
+The straightforward `fitBB` function provides an interface to non-linear or linear regression to fit one of three stomatal conductance models [@ballberry1987; @leuning1995; @medlyn2011].
+This yields estimates of g~1~ and (optionally ) g~0~, which are necessary inputs to the coupled leaf gas exchange model.
 
@@
 
-The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration. This is solved by the `Photosyn` function.
+The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration.
+This is solved by the `Photosyn` function.
+This flexible interface can be used to either a) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`, equivalent to `Aci(...)`), b) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf.
+Fig.
+2) or c) solve C~i~ from the coupled leaf gas exchange model (Eqs.
+1-3).
+
+- example response curves
+- energy balance (cite [@wang1998; @buckley2014]) - brief on details.
 
@@
 
-Full numerical optimization of the Cowan-Farquhar hypothesis. We have used this code in [@medlyn2011], and is very similar to that used by [@buckley2014].
+Full numerical optimization of the Cowan-Farquhar hypothesis.
+I have used this code in [@medlyn2011], and it is very similar to that presented (but not made publicly available) by [@buckley2014].
+It is visualized in Figure 4.
 
@@
 
-We need a publicly available, open source, set of tools to analyze these data which form the cornerstone of plant physiological ecology.
-Though alternative methods exist to fit A-Ci curves and other details of the implementations can see alternatives, the plantecophys package provides a platform where such methods comparisons can be made.
-The R language is now widely used, and many users can make changes to the code and recompile the package.
-We have included a set of functions that are key to ecophysiological research. There is certainly scope, and interest, to increase the array of tools included in the package.
+We need a publicly available, open source, set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
+Though alternative methods exist to fit A-C~i~ curves, the plantecophys package provides a platform where such methods comparisons can be made.
+As such I do not claim that `fitaci` is the only or best method available.
+The coupled leaf gas exchange model is widely used but no implementations are publicly available.
+The plantecophys package fills this void.
+
+The R language is now widely used, and more advanced users can make changes to the code and rebuild the package.
+Through the git version control system it is straightforward to include new changes into the central repository.
 
@@
 # Acknowledgements {.unnumbered}
-Thanks to John Drake for extensive testing.
 David Ellsworth, Belinda Medlyn, Martin de Kauwe are acknowledged for discussion on fitting A-Ci curves.
 Special thanks to Belinda Medlyn whose code in the MAESTRA model provided inspiration for an early version of the Photosyn function.
+Thanks to John Drake for testing.
 
@@
 
-```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
 figure1()
 ```
-** Figure 1**.
+
+** Figure 1**. Leaf photosynthesis - CO~2~ response curve as modelled with the FvC model.
+
+```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+suppressWarnings(figure2())
+```
+
+** Figure 2**.
 The intersection of the supply and demand curves of photosynthesis.
@@
 
-```{r fig.width=6, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure2()
+```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+figure3()
 ```
-**Figure 2**.
+
+**Figure 3**.
 Standard output from the `fitaci` function.
@@
 
+```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+figure4()
+```
+
+**Figure 4**.
+Visualization of the optimal model of stomatal conductance.
+Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration.
+In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda*E$ versus the C~i~, demonstrating that an optimum exists.
+The ~FARAO~ function finds this optimum numerically and calculates A~n~ and g~s~.
+The corresponding response of g~s~ to D is shown in panel (b).
+
 # References {.unnumbered}

### 2015-07-06 R/functions.R

@@ -7,6 +7,7 @@ to.pdf <- function(expr, filename, ..., verbose=TRUE) {
   on.exit(dev.off())
   eval.parent(substitute(expr))
 }
+
 to.eps <- function(expr, filename, ..., verbose=TRUE) {
   if(!file.exists(dirname(filename)))
     dir.create(dirname(filename), recursive=TRUE)
@@ -15,4 +16,15 @@ to.eps <- function(expr, filename, ..., verbose=TRUE) {
   postscript(filename, ...)
   on.exit(dev.off())
   eval.parent(substitute(expr))
+}
+
+# Simple function for placing labels on a figure.
+plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
+  u <- par()$usr
+  if(grepl("left",where))x <- u[1] + inset.x*(u[2]-u[1])
+  if(grepl("right",where))x <- u[2] - inset.x*(u[2]-u[1])
+  if(grepl("bottom",where))y <- u[3] + inset.y*(u[4]-u[3])
+  if(grepl("top",where))y <- u[4] - inset.y*(u[4]-u[3])
+  
+  text(x,y,txt,font=2,...)
 }

### 2015-07-06 figures.R

@@ -3,9 +3,35 @@ source("R/functions.R")
 library(plantecophys)
 
 
+figure1 <- function(){
+  
+  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
+  
+  ci <- seq(40, 1050, length=101)
+  acirun <- Aci(Ci=ci, Vcmax=65, Jmax=130, Rd=1.5)
+  
+  with(acirun, {
+    plot(Ci, Ac-Rd, type='l',col="red",lwd=2, ylim=c(0,40),
+         xlab=expression(italic(C)[i]~~(ppm)),
+         ylab=expression(italic(A)~~(mu*mol~m^-2~s^-1)),
+                    xlim=c(0,1050))
+    lines(Ci, Aj-Rd, col="blue", lwd=2)
+    lines(Ci, ALEAF, lwd=2)
+  })
+  
+  x <- plantecophys:::findCiTransition(Aci, Vcmax=65, Jmax=130, Rd=1.5)
+  points(x, Aci(x, Vcmax=65, Jmax=130, Rd=1.5)$ALEAF, pch=21, cex=1.4, bg="white")
+  
+  legend("topleft", c(expression(A[j]),expression(A[c]), expression(A[n]), "Transition point"),
+         lwd=c(2,2,2,-1), pch=c(-1,-1,-1,1), pt.cex=c(-1,-1,-1,1.4), col=c("blue","red","black", "black"),
+         bty='n')
+}
+
+
+
 
 # Sharkey figure
-figure1 <- function(){
+figure2 <- function(){
   ci <- seq(50, 800, length=101)
   
   acirun <- Aci(Ci=ci)
@@ -13,27 +39,25 @@ figure1 <- function(){
   par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
   with(acirun, plot(Ci, ALEAF, type='l',
                     ylim=c(0,20), xlim=c(0,800),
-                    col="blue",lwd=2,
+                    lwd=2,
                     xlab=expression(italic(C)[i]~~(ppm)),
-                    ylab=expression(italic(A)[leaf]~~(mu*mol~m^-2~s^-1))))
+                    ylab=expression(italic(A)[n]~~(mu*mol~m^-2~s^-1))))
   
   p <- Photosyn(VPD=2.5, g1=2)
   
-  gc <- p$GS / 1.6
[…]

### 2015-07-06 manuscript.Rmd
combine fig 1 and 2

@@
 
-```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
 figure1()
@@
 
-** Figure 1**. Leaf photosynthesis - CO~2~ response curve as modelled with the FvC model.
-
-```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-suppressWarnings(figure2())
-```
-
-** Figure 2**.
-The intersection of the supply and demand curves of photosynthesis.
+** Figure 1**.
+(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvC model.
+(b) The intersection of the supply and demand curves of photosynthesis.
 The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvC model) are known.
@@
 ```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure3()
+figure2()
 ```
 
-**Figure 3**.
+**Figure 2**.
 Standard output from the `fitaci` function.
@@
 ```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure4()
+figure3()
 ```
 
-**Figure 4**.
+**Figure 3**.
 Visualization of the optimal model of stomatal conductance.

### 2015-07-06 figures.R

@@ -5,7 +5,8 @@ library(plantecophys)
 
 figure1 <- function(){
   
-  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
+  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1,
+      mfrow=c(1,2))
   
   ci <- seq(40, 1050, length=101)
   acirun <- Aci(Ci=ci, Vcmax=65, Jmax=130, Rd=1.5)
@@ -25,15 +26,11 @@ figure1 <- function(){
   legend("topleft", c(expression(A[j]),expression(A[c]), expression(A[n]), "Transition point"),
          lwd=c(2,2,2,-1), pch=c(-1,-1,-1,1), pt.cex=c(-1,-1,-1,1.4), col=c("blue","red","black", "black"),
          bty='n')
-}
-
-
-
-
-# Sharkey figure
-figure2 <- function(){
-  ci <- seq(50, 800, length=101)
   
+  ci <- seq(50, 800, length=101)
+  plotlabel("(a)", "topleft")
+  
+  # PAnel b
   acirun <- Aci(Ci=ci)
   
   par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
@@ -49,6 +46,7 @@ figure2 <- function(){
   abline(gc*p$Ca, -gc,  lwd=2, lty=5)
   
   points(p$Ci, p$ALEAF, pch=21, bg="white", cex=1.4)
+  plotlabel("(b)", "topleft")
   
   legend("right", c("Demand function","Supply function","Operating point"),
          lwd=c(2,2,-1),pch=c(-1,-1,1), pt.cex=c(-1,-1,1.4),
@@ -56,8 +54,12 @@ figure2 <- function(){
 }
 
 
+
+
+
+
 # example output fitaci
-figure3 <- function(){
+figure2 <- function(){
   
   
   f <- fitaci(acidata1)
@@ -69,7 +71,7 @@ figure3 <- function(){
 
 
 # FARAO
-figure4 <- function(){
+figure3 <- function(){
   vpds <- seq(0.5,3.5, by=0.5)
   lam <- 0.002
   f <- plantecophys:::OPTfun

### 2015-07-06 manuscript.Rmd
45min of work on the paper.

@@
 ---
-title: plantecophys - an R package for analysis and modelling leaf gas exchange data
+title: plantecophys - an R package for analysing and modelling leaf gas exchange data
 author: ''
@@
 
-```{r echo=FALSE, message=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 # Two useful packages for printing numbers, accessing content of objects.
@@
 library(reporttools)
+
+# Tables in markdown
+library(ascii)
 
@@
 Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models.
-These models include the Farquhar-von Caemmerer-Berry model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO2 in the leaf.
+These models include the Farquhar-von Caemmerer-Berry model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf.
 The 'plantecophys' R package includes functions both for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data.
 Here I describe the functionality and implementation of the new package, and give examples on its use in plant ecophysiological research.
-I briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
+I briefly describe functions for modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, fitting the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
 The package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.
@@
 
-Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO2 and H~2~O has been collected.
+Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected.
 These data play a central role in physiological plant ecology, because rates of photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
@@
 This widely cited model is embedded in many global vegetation models (cite).
-The key prediction of the model is the response of photosynthesis to [CO~2~] in the intercellular airspaces (Ci).
-To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO2] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~).
+The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular CO~2~, C~i~).
+To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~).
 
@@
 The details of these functions and the temperature dependence of the parameters are described elsewhere [@medlyn2002].
-The use of C~i~ in these equations is routine, but actually it should be Cc, considering that CO2 in the chloroplast is not equal to that in the intercellular air space.
+We have here used C~i~ to denote [CO~2~] concentration in the leaf.
+Alternatively we may account for the diffusional limitation from intercellular airspaces to the chloroplast, by calculating C~c~ as,
+
+(2)
+$$C_c = C_i - A~n~/g~m~$$
+
+where g~m~ is the mesophyll conductance.
 
@@
 
-(2)
+(3)
 $$ A_n = g_s(C_a - C_i) $$
@@
 
-(3)
+(4)
 $$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$
@@
 The code is freely available (without restrictions), and managed with a version control system.
-I have developed the code base in my work on leaf and canopy modelling of photosynthesis and stomatal conductance, with many additions based on user requests [@medlyn2011; @barton2012; @medlyn2013; @duursma2013; @duursma2014].
+The package is the results of our work on leaf and canopy modelling of photosynthesis and stomatal conductance, with many additions based on user requests [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014].
 
@@
 ```{r results='asis', echo=FALSE, message=FALSE, warn=FALSE}
-f <- c("`fitaci`","`fitBB`","`FARAO`","`Photosyn`","`PhotosynEB`","`RHtoVPD` etc.")
+f <- c("`fitaci`","`fitBB`","`FARAO`","`Photosyn`","`PhotosynEB`","`AciC4`", "`RHtoVPD` etc.")
 d <- c("Fit, summarize, plot and simulate photosynthesis-[CO~2~] response curves (A-Ci curves)",
@@
 "Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
-"Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model",
+"Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvC model when either C~i~ or g~s~ is given as input",
 "Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
+"Simulates the dependence of C4 photosynthesis on the intercellular CO~2~ concentration",
 "Convert between commonly used units (relative humidity, vapour pressure deficit, dewpoint temperature)")
@@
 
-# library(knitr)
-# kable(df, format="markdown",
-#       caption="Table 1. Main functions in the plantecophys package. Each function has its own detailed help page  including example code.")
-
-library(ascii)
 suppressWarnings(print(ascii(df, caption="Table 1. Main functions in the plantecophys package.",
@@
 Instead of fitting the minimum function (Eq.
-1) it is actually far more practical to use a hyperbolic minimum, which avoids a discontinuity at the transition from A~c~ to A~j~ limitation.
+1), `fitaci~ uses the hyperbolic minimum, which avoids a discontinuity at the transition from A~c~ to A~j~ limitation.
+Finally, the user can provide an estimate of mesophyll conductance (g~m~) [following @ethier2004], in which case the fitted values of V~cmax~ and J~max~ can be interpreted as chloroplastic rates.
 
 \
-The function takes a dataframe as input, which includes A~n~, C~i~ and optionally leaf temperature (T~leaf~) and photosynthetically active radiation (PAR).
+The function takes a dataframe as input, which includes measurements of A~n~, C~i~ and optionally leaf temperature (T~leaf~) and photosynthetically active radiation (PAR).
 It then returns a special object which has a print and plot method.
+It is easily used like this:
 
@@
 
+A C4 model of leaf photosynthesis [@caemmerer2000] is also implemented (in `AciC4`), but at the moment it is only possible to fit the C3 model of leaf photosynthesis to A-Ci curves.
+
 ### Fitting stomatal conductance models
@@
 
-The intersection of the supply and demand curves of photosynthesis gives the operating intercellular CO~2~ concentration.
+The intersection of the supply and demand curves of photosynthesis (Fig.
+1b) gives the operating intercellular CO~2~ concentration.
 This is solved by the `Photosyn` function.
-This flexible interface can be used to either a) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`, equivalent to `Aci(...)`), b) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf.
+This flexible interface can be used to either 1) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`, equivalent to `Aci(...)`), 2) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf.
 Fig.
-2) or c) solve C~i~ from the coupled leaf gas exchange model (Eqs.
-1-3).
+1b) or c) solve C~i~ from the coupled leaf gas exchange model (Eqs.
+1,3,4).
 
-- example response curves
+To simulate photosynthesis across a range of environmental drivers, we can provide multiple values for either PAR, Tleaf, VPD, or CO2.
+In the following example we reduce J~max~ but hold V~cmax~ constant to look at the effect on the temperature optimum of A~n~.
+Note that VPD is held constant in this simple example.
+
+*because it is interesting to study this, since Jmax and Vcmax have different T responses, so we expect a different Jmax/Vcmax ratio to affect the T response of An*
+
+```
+# Set range of leaf temperature
+tleafs <- seq(10,40,by=0.5)
+
+# Simulate.
+run1 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=2*50, g1=3.5)
+run2 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=1.5*50, g1=3.5)
+```
+
+These results are then used to draw the results in Fig. 3.
+
+- leaf respiration follows a simple Q10 response.
+- temperature response functions built in following Medlyn et al. [@medlyn2002]
 - energy balance (cite [@wang1998; @buckley2014]) - brief on details.
@@
 
-**Figure 3**.
+**Figure 3**. `Photosyn` examples.
+
+```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+figure4()
+```
+
+**Figure 4**.
 Visualization of the optimal model of stomatal conductance.

### 2015-07-06 figures.R

@@ -70,8 +70,23 @@ figure2 <- function(){
 }
 
 
-# FARAO
 figure3 <- function(){
+  
+  # Set range of leaf temperature
+  tleafs <- seq(10,40,by=0.5)
+  
+  # Simulate.
+  run1 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=2*50, g1=3.5)
+  run2 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=1.5*50, g1=3.5)
+  
+  with(run1, plot(Tleaf, ALEAF, type='l', lwd=2,
+                  ylim=c(0,15), xlim=c(0,45)))
+  with(run2, lines(Tleaf, ALEAF, type='l', lty=5, lwd=2))
+}
+
+
+# FARAO
+figure4 <- function(){
   vpds <- seq(0.5,3.5, by=0.5)
   lam <- 0.002
   f <- plantecophys:::OPTfun

### 2015-07-07 manuscript.Rmd
First full mostly clean draft

@@
 
+# use git
+library(git2r)
+
 # figures
@@
 # render("manuscript.Rmd", "word_document", "manuscript.docx")
+
+# Find current SHA of plantecophys
+pkgrep <- repository("c:/repos/plantecophys")
+r <- reflog(pkgrep, verbose=FALSE)
+cur_sha <- substr(r[[1]]@sha, 1, 7)
 ```
@@
 
-Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected.
-These data play a central role in physiological plant ecology, because rates of photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
-Not only do leaf gas exchange data allow detailed studies of the underlying plant physiological controls, they are also used to parameterize an important component of land surface models used to predict global water and carbon cycling [e.g.
+Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected  [@long2003].
+These data play a central role in physiological plant ecology [@pearcy1989].
+Both photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
+Not only do leaf gas exchange data allow detailed studies of the underlying plant physiology, they are also used to parameterize an important component of land surface models used to predict global water and carbon cycling [e.g.
 @bonan2014; @dekauwe2015].
@@
 \
-The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting leaf gas exchange data (cites), by providing sensible and comparable metrics of the photosynthetic capacity, and accurately predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
-This widely cited model is embedded in many global vegetation models (cite).
+The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting leaf gas exchange data (cites), by providing sensible and comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
+This widely cited model is embedded in many process-based models of vegetation function.
 The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular CO~2~, C~i~).
+It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @medlyn2002].
 To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~).
@@
 \
-We do not repeat a detailed description of the FvC model here, as it has been described many times. But generally it is of the form,
+We do not repeat a detailed description of the FvC model here, as it has been described many times [e.g. @medlyn2002]. But generally it is of the form,
 
@@
 (2)
-$$C_c = C_i - A~n~/g~m~$$
+$$C_c = C_i - A_n/g_m$$
 
@@
 Moreover, sample sizes collected are often small due to time constraints in the field.
-A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that both prevents batch analysis and is arbitrary.
-Another method (http://leafweb.ornl.gov/) requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
+A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that is both both arbitrary and prevents batch analysis.
+Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
 Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]).
@@
 Through Eq.
-1, we have a dependency of photosynthesis on the availability of the substrate, Ci.
-To estimate Ci itself, we must have an estimate of the stomatal conductance to CO~2~.
+1, we have a dependency of photosynthesis on the availability of the substrate, C~i~.
+To estimate C~i~ itself, we need an estimate of the stomatal conductance to CO~2~.
 From Fick's law, we can relate A~n~ to g~s~ and C~i~ as,
@@
 3).
-This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991] is implemented in many process-based ecosystem models [@wangl1998; @bonan2014].
-This model allows prediction of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g1, Vcmax, Jmax, Rd).
+This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991] is implemented in many process-based ecosystem models [@wang1998; @bonan2014].
+This model allows prediction of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~).
 
@@
 Fitting the FvC model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
-The coupled leaf gas exchange model is not straightforward to implement, and I do not know of an open-source implementation that is publicly available.
+The coupled leaf gas exchange model is not straightforward to implement, and I do not know of any standalone open-source implementations.
 I here describe the plantecophys package, implemented in the R language [@rfoundation].
 The code is freely available (without restrictions), and managed with a version control system.
-The package is the results of our work on leaf and canopy modelling of photosynthesis and stomatal conductance, with many additions based on user requests [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014].
+The package is the results of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014], with many additions based on user requests.
 
@@
 "Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
-"Simulate leaf photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvC model when either C~i~ or g~s~ is given as input",
+"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvC model when either C~i~ or g~s~ is given as input. For C4 photosynthesis, see `AciC4`",
 "Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
@@
 The repository includes an issue tracker, where users can suggest changes or report bugs.
+This paper describes version `r packageVersion("plantecophys")` (git SHA `r cur_sha`).
 
@@
 
+\
 A C4 model of leaf photosynthesis [@caemmerer2000] is also implemented (in `AciC4`), but at the moment it is only possible to fit the C3 model of leaf photosynthesis to A-Ci curves.
@@
 
-To simulate photosynthesis across a range of environmental drivers, we can provide multiple values for either PAR, Tleaf, VPD, or CO2.
-In the following example we reduce J~max~ but hold V~cmax~ constant to look at the effect on the temperature optimum of A~n~.
-Note that VPD is held constant in this simple example.
-
-*because it is interesting to study this, since Jmax and Vcmax have different T responses, so we expect a different Jmax/Vcmax ratio to affect the T response of An*
+To demonstrate the use of the coupled gas exchange model, I visualize the temperature response of A~n~ when both T~leaf~ and D are varying.
+In field conditions, D is always strongly positively related to T~leaf~.
+The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014].
+In the FvC model, J~max~, V~cmax~ and leaf respiration (R~d~) all depend on T~leaf~.
+The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvC model [following @medlyn2002].
+Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D [calculated following @duursma2014].
+Figure 3b shows the output from `Photosyn`, which can be created with the following command.
+Note that in this example the default values of many parameters are used in the call to `Photosyn`, but all of these can be changed by the user.
 
@@
 # Set range of leaf temperature
-tleafs <- seq(10,40,by=0.5)
+tleafs <- seq(5, 40, by=5)
+
+# Define D as a function of Tleaf
+vpdfun <- function(tair)0.000605*tair^2.39
 
 # Simulate.
-run1 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=2*50, g1=3.5)
-run2 <- Photosyn(Tleaf = tleafs, Vcmax=50, Jmax=1.5*50, g1=3.5)
+run1 <- Photosyn(Tleaf = tleafs, VPD=vpdfun(tleafs))
+
+# Plot
+with(run1, plot(Tleaf, ALEAF))
+
 ```
 
-These results are then used to draw the results in Fig. 3.
-
-- leaf respiration follows a simple Q10 response.
-- temperature response functions built in following Medlyn et al. [@medlyn2002]
-- energy balance (cite [@wang1998; @buckley2014]) - brief on details.
+The `Photosyn` function assumes that the boundary layer conductance (g~bl~) is high compared to g~s~, so that T~leaf~ is close to T~air~.
+As an alternative, the `PhotosynEB` function calculates T~leaf~ from the leaf energy balance.
+Transpiration is calculated with the Penman-Monteith equation, which accounts for boundary layer effects.
+The details of `PhotosynEB` are not described here (see the documentation), because it is very similar to other implementations [@wang1998; @buckley2014].
 
@@
 
-Full numerical optimization of the Cowan-Farquhar hypothesis.
-I have used this code in [@medlyn2011], and it is very similar to that presented (but not made publicly available) by [@buckley2014].
-It is visualized in Figure 4.
+The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar hypothesis.
+It find the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates A~n~ as well as g~s~.
+This implementation was used by Medlyn et al.
+[@medlyn2011] to compare a simplified model of optimal stomatal conductance to the full numerical solution.
+In Fig.
+4a, I have calculated $A - \lambda E$ across a range of C~i~ values, and for different values of D.
+The `FARAO` function calculates the optima of these curves, which can for example be used to study the response of stomatal conductance to D (Fig.
+4b).
+
+\
+Optionally, the `FARAO` function accounts for the presence of a leaf boundary layer (when `energybalance=TRUE`).
+In that case it uses `PhotosynEB` (see description above) to calculate A~n~ and E, and solves for T~leaf~.
+A very similar method was employed by Buckley et al.
+[@buckley2014], who demonstrated that when a boundary layer is present, frequently an optimal g~s~ cannot be found.
 
@@
 
-We need a publicly available, open source, set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
-Though alternative methods exist to fit A-C~i~ curves, the plantecophys package provides a platform where such methods comparisons can be made.
-As such I do not claim that `fitaci` is the only or best method available.
-The coupled leaf gas exchange model is widely used but no implementations are publicly available.
-The plantecophys package fills this void.
-
-The R language is now widely used, and more advanced users can make changes to the code and rebuild the package.
-Through the git version control system it is straightforward to include new changes into the central repository.
+We need an open source set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
+The plantecophys R package contains a useful set of tools to perform standard, and more advanced, analyses.
+The open source framework combined with version control allows further development of the code.
 
@@
 
-```{r fig.width=6, fig.height=5, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
+```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
 figure2()
@@
 Standard output from the `fitaci` function.
-A~net~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration.
-Symbols are measurements, the line the fitted Farquhar model of photosynthesis.
+A~n~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration.
+Symbols are measurements, the black line the fitted FvC model of photosynthesis.
+Colored lines indicate the two photosynthesis rates in the FvC model.
 
@@
 
-**Figure 3**. `Photosyn` examples.
+**Figure 3**.
+Response of A~n~ and C~i~ to combined changes in T~leaf~ and D.
+(a) Lines are A-C~i~ curves simulated at a range of values for T~leaf~.
+Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014]).
+Note that as T~leaf~ and D increase, C~i~ decreases.
+(b) The corresponding temperature optimum of A~n~.
+Symbols are the same as in panel (a) but plotted against T~leaf~.
 

### 2015-07-07 R/functions.R

@@ -28,3 +28,39 @@ plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
   
   text(x,y,txt,font=2,...)
 }
+
+
+#'@title Add a line to a plot
+#'@description As \code{abline}, but with \code{from} and \code{to} arguments. 
+#'If a fitted linear regression model is used as asn argument, it uses the min and max values of the data used to fit the model.
+#'@param a Intercept (optional)
+#'@param b Slope (optional)
+#'@param reg A fitted linear regression model (output of \code{\link{lm}}).
+#'@param from Draw from this X value
+#'@param to Draw to this x value
+#'@param \dots Further parameters passed to \code{\link{segments}}
+#'@export
+ablinepiece <- function(a=NULL,b=NULL,reg=NULL,from=NULL,to=NULL,...){
+  
+  # Borrowed from abline
+  if (!is.null(reg)) a <- reg
+  
+  if (!is.null(a) && is.list(a)) {
+    temp <- as.vector(coefficients(a))
+    from <- min(a$model[,2], na.rm=TRUE)
+    to <- max(a$model[,2], na.rm=TRUE)
+    
+    if (length(temp) == 1) {
+      a <- 0
+      b <- temp
+    }
+    else {
+      a <- temp[1]
+      b <- temp[2]
+    }
+  }
+  
+  segments(x0=from,x1=to,
+           y0=a+from*b,y1=a+to*b,...)
+  
+}

### 2015-07-07 figures.R

@@ -25,10 +25,10 @@ figure1 <- function(){
   
   legend("topleft", c(expression(A[j]),expression(A[c]), expression(A[n]), "Transition point"),
          lwd=c(2,2,2,-1), pch=c(-1,-1,-1,1), pt.cex=c(-1,-1,-1,1.4), col=c("blue","red","black", "black"),
-         bty='n')
+         bty='n', cex=0.9)
   
   ci <- seq(50, 800, length=101)
-  plotlabel("(a)", "topleft")
+  plotlabel("(a)", "bottomright")
   
   # PAnel b
   acirun <- Aci(Ci=ci)
@@ -45,12 +45,12 @@ figure1 <- function(){
   gc <- p$GS / 1.57
   abline(gc*p$Ca, -gc,  lwd=2, lty=5)
   
-  points(p$Ci, p$ALEAF, pch=21, bg="white", cex=1.4)
-  plotlabel("(b)", "topleft")
+  points(p$Ci, p$ALEAF, pch=22, bg="white", cex=1.4)
+  plotlabel("(b)", "bottomright")
   
-  legend("right", c("Demand function","Supply function","Operating point"),
-         lwd=c(2,2,-1),pch=c(-1,-1,1), pt.cex=c(-1,-1,1.4),
-         inset=0.01, cex=1.1, bty='n', lty=c(1,5,-1))
+  legend("right", c("Demand","Supply","Solution"),
+         lwd=c(2,2,-1),pch=c(-1,-1,0), pt.cex=c(-1,-1,1.4),
+         inset=0.01, cex=0.8, bty='n', lty=c(1,5,-1))
 }
 
 
@@ -61,8 +61,8 @@ figure1 <- function(){
 # example output fitaci
 figure2 <- function(){
   
-  
-  f <- fitaci(acidata1)
+  acidata1$PPFD <- 1800
+  f <- suppressWarnings(fitaci(acidata1))
   
   par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1)
   plot(f)
@@ -70,19 +70,57 @@ figure2 <- function(){
 }
 
 
+
 figure3 <- function(){
   
-  # Set range of leaf temperature
-  tleafs <- seq(10,40,by=0.5)
+  #T response 
+  vpdfun <- function(tair)0.000605*tair^2.39
   
[…]

### 2015-07-07 manuscript.Rmd
Minor fixes to refs

@@
 Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models.
-These models include the Farquhar-von Caemmerer-Berry model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf.
+These models include the Farquhar-von Caemmerer-Berry (FvCB) model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf.
 The 'plantecophys' R package includes functions both for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data.
 Here I describe the functionality and implementation of the new package, and give examples on its use in plant ecophysiological research.
-I briefly describe functions for modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, fitting the Farquhar model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
-The package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.
+I briefly describe functions for modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, fitting the FvCB model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
+This open-source package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.
 
@@
 \
-The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting leaf gas exchange data (cites), by providing sensible and comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
-This widely cited model is embedded in many process-based models of vegetation function.
+The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting and modelling leaf gas exchange, by providing comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
+This widely cited model is embedded in many process-based models of vegetation function [@sellers1997; @bonan2014].
 The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular CO~2~, C~i~).
@@
 This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991] is implemented in many process-based ecosystem models [@wang1998; @bonan2014].
-This model allows prediction of A, g_s and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~).
+This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~).
 
 \
-Despite the widespread use of the FvC model and the coupled leaf gas exchange model, tools to analyze the data are scattered and subject to little standardization.
-Fitting the FvC model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
+Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyze the data are scattered and subject to little standardization.
+Fitting the FvCB model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
 The coupled leaf gas exchange model is not straightforward to implement, and I do not know of any standalone open-source implementations.
@@
 "Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
-"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvC model when either C~i~ or g~s~ is given as input. For C4 photosynthesis, see `AciC4`",
+"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvCB model when either C~i~ or g~s~ is given as input. For C4 photosynthesis, see `AciC4`",
 "Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
@@
 
-The `fitaci` function fits the FvC model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors.
+The `fitaci` function fits the FvCB model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors.
 Optionally R~d~ can be provided as a known value, otherwise it is estimated from the A-Ci curve.
@@
 ```
-# Fit FvC model
+# Fit FvCB model
 f <- fitaci(mydata)
@@
 The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014].
-In the FvC model, J~max~, V~cmax~ and leaf respiration (R~d~) all depend on T~leaf~.
-The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvC model [following @medlyn2002].
+In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) all depend on T~leaf~.
+The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model [following @medlyn2002].
 Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D [calculated following @duursma2014].
@@
 ** Figure 1**.
-(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvC model.
+(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model.
 (b) The intersection of the supply and demand curves of photosynthesis.
-The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvC model) are known.
+The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
 
@@
 A~n~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration.
-Symbols are measurements, the black line the fitted FvC model of photosynthesis.
-Colored lines indicate the two photosynthesis rates in the FvC model.
+Symbols are measurements, the black line the fitted FvCB model of photosynthesis.
+Colored lines indicate the two photosynthesis rates in the FvCB model.
 

### 2015-07-07 manuscript.Rmd
Formatting requirements for PLOS ONE; reference fixes

@@
 T: +61(0)422096908
+
+*Short title*: The plantecophys R package
 
@@
 
+# output
+if(!dir.exists("output"))dir.create("output")
+
 # Do calculations here, create objects that we can use throughout the paper.
@@
 # Set Flag to include figures in text or not.
-includeFigs <- TRUE
+includeFigs <- FALSE
 
@@
 # Find current SHA of plantecophys
-pkgrep <- repository("c:/repos/plantecophys")
+rep_pkg <- "c:/repos/plantecophys"
+if(file.exists(rep_pkg)){
+  pkgrep <- repository(rep_pkg)
   r <- reflog(pkgrep, verbose=FALSE)
   cur_sha <- substr(r[[1]]@sha, 1, 7)
+} else {
+  cur_sha <- "???"
+}
+
+# Make PDFs
+to.pdf(figure1(), "output/figure1.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure2(), "output/figure2.pdf", width=4, height=4, verbose=FALSE)
+to.pdf(figure3(), "output/figure3.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure4(), "output/figure4.pdf", width=8, height=4, verbose=FALSE)
 ```
@@
 
+Keywords: photosynthesis, stomatal conductance, portable gas exchange instrument, process-based model, software
+
 # Introduction
 
-Since the advent of portable gas exchange instruments, a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected  [@long2003].
-These data play a central role in physiological plant ecology [@pearcy1989].
+Since the advent of portable gas exchange instruments [@mcdermitt1989; @leuning1989], a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected  [@long2003].
+These data play a central role in physiological plant ecology [@pearcey1989].
 Both photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
@@
 
+\
 (1)
@@
 
+\
 (2)
@@
 Moreover, sample sizes collected are often small due to time constraints in the field.
-A widely used published method requires the user to specify the transition of Vcmax to Jmax limitation [@sharkey2007], a process that is both both arbitrary and prevents batch analysis.
-Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process [@gu2010].
+A widely used published method requires the user to specify the transition of V~cmax~ to J~max~ limitation [@sharkey2007], a process that is both arbitrary and prevents batch analysis.
+Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process [following @gu2010].
 Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]).
@@
 
+\
 (3)
@@
 where g~s~ is the conductance to CO~2~.
-We now have two equations for the photosynthesis rate.
+We now have two equations for A~n~.
 The 'demand function' is given by Eq.
 1, and the 'supply function' in Eq.
-2.
+3.
 At steady state these two equation should be equal, which can be graphically shown as in Fig.
@@
 \
-Because g~s~ itself responds to environmental drivers, we must then estimate this dependence with another model.
-The most successful, though empirical, g~s~ models is the Ball-Berry [@ballberry1987] class of models.
-This model posits an entirely empirical equation that describes the response of g~s~ to air humidity, CO~2~ and the rate of photosynthesis.
+**Figure 1**.
+(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model.
+(b) The intersection of the supply and demand curves of photosynthesis.
+The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
+
+\
+Because g~s~ itself responds to environmental drivers, another expression is needed to end up with a fully coupled model of leaf gas exchange.
+The most successful, though empirical, g~s~ model is the Ball-Berry [@ballberry1987] class of models.
+This model posits an entirely empirical equation that describes the response of g~s~ to air humidity, CO~2~ and A~n~.
 This way, effects of leaf temperature and PPFD - both of which are known to affect g~s~ - are modelled through the dependency of A~n~ on these drivers.
-Thus the third equation is,
+A general form of the Ball-Berry model is,
 
+\
 (4)
@@
 \
-where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters.
+where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters, and f(D) a function that describes the response to the vapour pressure deficit (D, [@leuning1995]) or relative humidity [@ballberry1987].
 An alternative approach to modelling g~s~ is through the hypothesis that stomata act optimally in the sense that they maximize photosynthesis while minimizing water loss.
@@
 Medlyn et al.
-[@medlyn2011] showed that the optimality hypothesis, when coupled to the FvC model, leads to an expression analogous to the Ball-Berry type models (Eq.
-3), but with a different D response function (f(D) in Eq.
-3) than the original Ball-Berry model, or yet another version introduced by Leuning [@leuning1995].
+[@medlyn2011] showed that the optimality hypothesis, when coupled to the FvCB model, leads to an expression analogous to the Ball-Berry type models (Eq.
+4), but with a different D response function (f(D) in Eq.
+3) than the original Ball-Berry model.
 
@@
 3).
-This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991] is implemented in many process-based ecosystem models [@wang1998; @bonan2014].
+This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991; @leuning1995] is implemented in many process-based ecosystem and global land surface models [@wang1998; @bonan2014; @duursma2012; @sellers1997].
 This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~).
@@
 The code is freely available (without restrictions), and managed with a version control system.
-The package is the results of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014], with many additions based on user requests.
+The package is the result of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014], with many additions based on user requests.
 
@@
 
-The 'plantecophys' package is implemented in native R, and has no dependencies on other packages.
+The 'plantecophys' package is implemented in R, has no dependencies on other packages, and does not require compilation (i.e.
+it is written in native R only).
 As such it builds easily, and is highly portable.
@@
 
-### Fitting A-Ci curves
+### Fitting A-C~i~ curves
 
 The `fitaci` function fits the FvCB model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors.
-Optionally R~d~ can be provided as a known value, otherwise it is estimated from the A-Ci curve.
+Optionally R~d~ can be provided as a known value, otherwise it is estimated from the A-C~i~ curve.
 The user does not have to provide the transition point (see Fig.
 1), as this is estimated by `fitaci` automatically.
-It is however an option to fix the transition point, which may be helpful to check whether the best fit was achieved.
+It is however an option to fix the transition point (via the `citransition` argument), which may be helpful to check whether the best fit was achieved.
 Standard errors are estimated with the usual methods for non-linear regression (`nls` function in base R).
 Instead of fitting the minimum function (Eq.
-1), `fitaci~ uses the hyperbolic minimum, which avoids a discontinuity at the transition from A~c~ to A~j~ limitation.
+1), `fitaci` uses the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity.
 Finally, the user can provide an estimate of mesophyll conductance (g~m~) [following @ethier2004], in which case the fitted values of V~cmax~ and J~max~ can be interpreted as chloroplastic rates.
+
+\
+In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) (and other parameters like $\Gamma^{*}$, K~c~ and K~o~) all depend non-linearly on T~leaf~.
+The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model [following @medlyn2002].
 
@@
 2.
-Additionally, `fitacis` can be sued to fit many curves at once, for example one for each species or site in a dataset.
+Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.
 The function `Aci` simulates the A-Ci curve, that is, it returns A~n~ when C~i~ is known.
+
+\
+**Figure 2**.
+Standard output from the `fitaci` function.
+A~n~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration.
+Symbols are measurements, the black line the fitted FvCB model of photosynthesis.
+Colored lines indicate the two photosynthesis rates in the FvCB model.
 
@@
 Fig.
-1b) or c) solve C~i~ from the coupled leaf gas exchange model (Eqs.
+1b) or c) solve for C~i~ from the coupled leaf gas exchange model (Eqs.
 1,3,4).
 
+\
 To demonstrate the use of the coupled gas exchange model, I visualize the temperature response of A~n~ when both T~leaf~ and D are varying.
@@
 The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014].
-In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) all depend on T~leaf~.
-The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model [following @medlyn2002].
 Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D [calculated following @duursma2014].
-Figure 3b shows the output from `Photosyn`, which can be created with the following command.
-Note that in this example the default values of many parameters are used in the call to `Photosyn`, but all of these can be changed by the user.
+Figure 3b shows the output from `Photosyn`, which can be created with the following example Note that in this example the default values of many parameters (e.g.
+J~max~, g~1~) are used in the call to `Photosyn`, but all of these can be set by the user.
+
+\
+**Figure 3**.
+Response of A~n~ and C~i~ to combined changes in T~leaf~ and D.
+(a) Lines are A-C~i~ curves simulated at a range of values for T~leaf~.
+Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014]).
+Note that as T~leaf~ and D increase, C~i~ decreases.
+(b) The corresponding temperature optimum of A~n~.
+Symbols are the same as in panel (a) but plotted against T~leaf~.
 
@@
 
+\
 The `Photosyn` function assumes that the boundary layer conductance (g~bl~) is high compared to g~s~, so that T~leaf~ is close to T~air~.
@@
 4b).
+
+\
+**Figure 4**.
+Visualization of the optimal model of stomatal conductance.
+Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration.
+In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda*E$ versus the C~i~, demonstrating that an optimum exists.
+The ~FARAO~ function finds this optimum numerically and calculates A~n~ and g~s~.
+The corresponding response of g~s~ to D is shown in panel (b).
 
@@
 
-# Acknowledgements {.unnumbered}
+# Acknowledgments {.unnumbered}
 David Ellsworth, Belinda Medlyn, Martin de Kauwe are acknowledged for discussion on fitting A-Ci curves.
@@
 
-# Figures {.unnumbered}
-
-```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure1()
-```
-
-** Figure 1**.
-(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model.
-(b) The intersection of the supply and demand curves of photosynthesis.
-The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
-
-```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure2()
-```
-
-**Figure 2**.
-Standard output from the `fitaci` function.
-A~n~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration.
-Symbols are measurements, the black line the fitted FvCB model of photosynthesis.
-Colored lines indicate the two photosynthesis rates in the FvCB model.
-
-```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure3()
-```
-
-**Figure 3**.
-Response of A~n~ and C~i~ to combined changes in T~leaf~ and D.
-(a) Lines are A-C~i~ curves simulated at a range of values for T~leaf~.
-Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014]).
-Note that as T~leaf~ and D increase, C~i~ decreases.
-(b) The corresponding temperature optimum of A~n~.
-Symbols are the same as in panel (a) but plotted against T~leaf~.
-
-```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs, warn=FALSE}
-figure4()
-```
-
-**Figure 4**.
-Visualization of the optimal model of stomatal conductance.
-Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration.
-In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda*E$ versus the C~i~, demonstrating that an optimum exists.
-The ~FARAO~ function finds this optimum numerically and calculates A~n~ and g~s~.
-The corresponding response of g~s~ to D is shown in panel (b).
-
 # References {.unnumbered}

### 2015-07-07 figures.R

@@ -64,8 +64,8 @@ figure2 <- function(){
   acidata1$PPFD <- 1800
   f <- suppressWarnings(fitaci(acidata1))
   
-  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1)
-  plot(f)
+  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1, xaxs="i", yaxs="i")
+  plot(f, xlim=c(0,1500), ylim=c(0,30))
   
 }
 
@@ -90,7 +90,7 @@ figure3 <- function(){
     
   }
   tairs <- seq(5,40, by=5)
-  Cols <- heat.colors(length(tairs)+3)
+  Cols <- heat.colors(length(tairs)+4)
   p <- Photosyn(Tleaf=tairs, VPD=vpdfun(tairs))
   
   
@@ -102,13 +102,13 @@ figure3 <- function(){
                                   cex=1.2,  bg=Cols[i])
   box()
   plotlabel("(a)","topleft")
-  for(i in 1:(nrow(p)-1)){
-    arrows(x0=p$Ci[i], x1=p$Ci[i+1], y0=p$ALEAF[i], y1=p$ALEAF[i+1],
-           col=Cols[i], length=0.15)
-  }
+#   for(i in 1:(nrow(p)-1)){
+#     arrows(x0=p$Ci[i], x1=p$Ci[i+1], y0=p$ALEAF[i], y1=p$ALEAF[i+1],
+#            col=Cols[i], length=0.15)
+#   }
   legend("left", c(expression(T[leaf] == 5),
                       expression(T[leaf] == 40)),
-         lty=1, lwd=2, col=Cols[c(1,length(Cols))], bty='n', cex=0.8)
+         lty=1, lwd=2, col=Cols[c(1,length(tairs))], bty='n', cex=0.8)
   
   
   with(p, plot(Tleaf, ALEAF,

### 2015-07-09 manuscript.Rmd
Final polishing; as submitted to PLOS ONE 2015-07-09

@@
 ---
-title: plantecophys - an R package for analysing and modelling leaf gas exchange data
+title: "plantecophys - an R package for analysing and modelling leaf gas exchange data"
 author: ''
@@
 
-Here I present the R package 'plantecophys', a toolkit to analyse and model leaf gas exchange data, a cornerstone of plant ecophysiological research.
+Here I present the R package 'plantecophys', a toolkit to analyse and model leaf gas exchange data.
 Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models.
 These models include the Farquhar-von Caemmerer-Berry (FvCB) model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf.
-The 'plantecophys' R package includes functions both for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data.
-Here I describe the functionality and implementation of the new package, and give examples on its use in plant ecophysiological research.
-I briefly describe functions for modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, fitting the FvCB model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-Ci curves'), modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
+The 'plantecophys' R package includes functions  for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data.
+Here I describe the functionality and implementation of the new package, and give some examples of its use.
+I briefly describe functions for fitting the FvCB model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-C~i~ curves'), fitting Ball-Berry type models, modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.
 This open-source package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.
 
+\
 Keywords: photosynthesis, stomatal conductance, portable gas exchange instrument, process-based model, software
@@
 Since the advent of portable gas exchange instruments [@mcdermitt1989; @leuning1989], a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected  [@long2003].
-These data play a central role in physiological plant ecology [@pearcey1989].
-Both photosynthesis and transpiration vary widely between plant species, and fluctuate rapidly in response to environmental drivers such as light, humidity and temperature.
-Not only do leaf gas exchange data allow detailed studies of the underlying plant physiology, they are also used to parameterize an important component of land surface models used to predict global water and carbon cycling [e.g.
+These data play a central role in physiological plant ecology [@pearcey1989], to better understand and quantify inter-specific differences in photosynthesis and transpiration, and to quantify and model the rapid response to changes in environmental drivers such as light, humidity and temperature.
+Not only do leaf gas exchange data allow detailed studies of the underlying plant physiology, they are also used to parameterize an important component of process-based models of vegetation function used to predict global water and carbon cycling [e.g.
 @bonan2014; @dekauwe2015].
@@
 \
-The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvC model') is widely used in interpreting and modelling leaf gas exchange, by providing comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
+The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvCB model') is widely used in interpreting and modelling leaf gas exchange, by providing comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~).
 This widely cited model is embedded in many process-based models of vegetation function [@sellers1997; @bonan2014].
-The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular CO~2~, C~i~).
+The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular [CO~2~], C~i~).
 It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @medlyn2002].
-To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~).
+To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~) [@wullschleger1993].
 
 \
-We do not repeat a detailed description of the FvC model here, as it has been described many times [e.g. @medlyn2002]. But generally it is of the form,
+We do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @medlyn2002]. But generally it is of the form,
 
@@
 \
-where A~c~ is the gross photosynthesis rate when Rubisco activity is limiting, A~j~ when RuBP-regeneration is limiting, and R~d~ the rate of dark respiration.
-A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i\Gamma^∗)/(k_2 + C_i)$, where $\Gamma^∗$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
-The details of these functions and the temperature dependence of the parameters are described elsewhere [@medlyn2002].
+where A~n~ is the net rate of CO~2~ assimulation, A~c~ is the gross photosynthesis rate when Rubisco activity is limiting, A~j~ when RuBP-regeneration is limiting, and R~d~ the rate of dark respiration (see Fig.
+1a).
+A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i{\Gamma}^{∗})/(k_2 + C_i)$, where ${\Gamma}^{∗}$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
+The details of these functions and the temperature dependence of the various parameters are described elsewhere [@medlyn2002].
 We have here used C~i~ to denote [CO~2~] concentration in the leaf.
-Alternatively we may account for the diffusional limitation from intercellular airspaces to the chloroplast, by calculating C~c~ as,
+Alternatively we may account for the diffusional limitation from intercellular airspaces to the chloroplast [@ethier2004], by calculating C~c~ as,
 
@@
 \
-Fitting the model to data requires some finesse because the model is not a single non-linear equation, but instead a minimum function of two non-linear equations that is sometimes difficult to fit.
+**Figure 1**.
+(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model.
+(b) The intersection of the supply and demand curves of photosynthesis.
+The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
+
+\
+Fitting the model to data requires some finesse because the model is a minimum function of two non-linear equations that is sometimes difficult to fit.
 Moreover, sample sizes collected are often small due to time constraints in the field.
 A widely used published method requires the user to specify the transition of V~cmax~ to J~max~ limitation [@sharkey2007], a process that is both arbitrary and prevents batch analysis.
-Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process [following @gu2010].
+Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process (following [@gu2010]).
 Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]).
+What is missing is an open-source tool that can be used for reproducible and transparent analysis of A-C~i~ curves.
 
@@
 At steady state these two equation should be equal, which can be graphically shown as in Fig.
-1 (cf.
+1b (cf.
 [@farquhar1982]).
@@
 \
-**Figure 1**.
-(a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model.
-(b) The intersection of the supply and demand curves of photosynthesis.
-The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
-
-\
 Because g~s~ itself responds to environmental drivers, another expression is needed to end up with a fully coupled model of leaf gas exchange.
-The most successful, though empirical, g~s~ model is the Ball-Berry [@ballberry1987] class of models.
+The most widely-used, though empirical, g~s~ model is the Ball-Berry [@ballberry1987] class of models.
 This model posits an entirely empirical equation that describes the response of g~s~ to air humidity, CO~2~ and A~n~.
@@
 This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991; @leuning1995] is implemented in many process-based ecosystem and global land surface models [@wang1998; @bonan2014; @duursma2012; @sellers1997].
-This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~).
+This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~, and their temperature dependencies).
 
 \
-Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyze the data are scattered and subject to little standardization.
+Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyze data and perform simulations are scattered and subject to little standardization.
 Fitting the FvCB model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
@@
 The code is freely available (without restrictions), and managed with a version control system.
-The package is the result of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014], with many additions based on user requests.
+The package is the result of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014; @lin2015], with many additions based on user requests.
 
@@
 
-The main functions in the package are summarized in Table 1.
+The main tools included in the plantecophys package are to a) fit A-C~i~ curves, b) fit Ball-Berry type models, c) simulate from the coupled leaf gas exchange model and d) calculate the optimal stomatal conductance.
+A list of the key functions in the package are summarized in Table 1.
 
@@
 The `fitaci` function fits the FvCB model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors.
-Optionally R~d~ can be provided as a known value, otherwise it is estimated from the A-C~i~ curve.
+Instead of fitting the minimum function (Eq.
+1), `fitaci` fits the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity.
+The fit is achieved via standard non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R).
+The `fitaci` function includes methods to estimate appropriate starting values from the data, and attempts the fits along a wide range of possible starting values.
+Optionally, R~d~ can be provided as a known value, otherwise it is estimated from the A-C~i~ curve.
 The user does not have to provide the transition point (see Fig.
@@
 It is however an option to fix the transition point (via the `citransition` argument), which may be helpful to check whether the best fit was achieved.
-Standard errors are estimated with the usual methods for non-linear regression (`nls` function in base R).
-Instead of fitting the minimum function (Eq.
-1), `fitaci` uses the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity.
-Finally, the user can provide an estimate of mesophyll conductance (g~m~) [following @ethier2004], in which case the fitted values of V~cmax~ and J~max~ can be interpreted as chloroplastic rates.
+Finally, the user can provide an estimate of mesophyll conductance (g~m~) (following [@ethier2004]), in which case the fitted values of V~cmax~ and J~max~ can be interpreted as chloroplastic rates.
 
@@
 In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) (and other parameters like $\Gamma^{*}$, K~c~ and K~o~) all depend non-linearly on T~leaf~.
-The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model [following @medlyn2002].
+The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model (following [@medlyn2002]).
+Optionally, measured K~m~ and $\Gamma^{*}$ can be provided instead.
 
@@
 
+# Print a summary with coefficients and more
+f
+
 # Make standard plot
@@
 \
-The output of the above example is shown in Fig.
-2.
-Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.
-The function `Aci` simulates the A-Ci curve, that is, it returns A~n~ when C~i~ is known.
+The output of the above example is shown in Fig. 2. Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.
 
@@
 \
-A C4 model of leaf photosynthesis [@caemmerer2000] is also implemented (in `AciC4`), but at the moment it is only possible to fit the C3 model of leaf photosynthesis to A-Ci curves.
+A C4 model of leaf photosynthesis [@caemmerer2000] is also implemented (in `AciC4`), but at the moment it is only possible to fit the C3 model of leaf photosynthesis to A-C~i~ curves.
 
@@
 This yields estimates of g~1~ and (optionally ) g~0~, which are necessary inputs to the coupled leaf gas exchange model.
+Note that, unlike the description in this paper, the user must provide stomatal conductance to H~2~O (not CO~2~) as input to the fitting process, which is the standard output of portable gas exchange instruments.
 
@@
 The intersection of the supply and demand curves of photosynthesis (Fig.
-1b) gives the operating intercellular CO~2~ concentration.
+1b) gives the steady-state intercellular CO~2~ concentration.
 This is solved by the `Photosyn` function.
@@
 The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014].
-Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D [calculated following @duursma2014].
-Figure 3b shows the output from `Photosyn`, which can be created with the following example Note that in this example the default values of many parameters (e.g.
+Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D (calculated following [@duursma2014]).
+Figure 3b shows the output from `Photosyn`, which can be created with the following example.
+Note that in this example the default values of many parameters (e.g.
 J~max~, g~1~) are used in the call to `Photosyn`, but all of these can be set by the user.
@@
 
-# Plot
+# Plot (produces Figure 3b minus the formatting)
 with(run1, plot(Tleaf, ALEAF))
@@
 Transpiration is calculated with the Penman-Monteith equation, which accounts for boundary layer effects.
-The details of `PhotosynEB` are not described here (see the documentation), because it is very similar to other implementations [@wang1998; @buckley2014].
+The details of `PhotosynEB` are not described here (we refer to the documentation), because it is very similar to other implementations [@wang1998; @buckley2014].
 
@@
 The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar hypothesis.
-It find the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates A~n~ as well as g~s~.
+It finds the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates corresponding A~n~ and g~s~.
 This implementation was used by Medlyn et al.
@@
 Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration.
-In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda*E$ versus the C~i~, demonstrating that an optimum exists.
-The ~FARAO~ function finds this optimum numerically and calculates A~n~ and g~s~.
+In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda E$ versus the C~i~, demonstrating that an optimum exists.
+The `FARAO` function finds this optimum numerically and calculates A~n~ and g~s~.
 The corresponding response of g~s~ to D is shown in panel (b).
@@
 We need an open source set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
-The plantecophys R package contains a useful set of tools to perform standard, and more advanced, analyses.
+At the moment there are no publicly available tools to fit A-C~i~ curves that can be used as part of a reproducable workflow.
+The plantecophys R package is implemented in widely used language for data analysis.
+The package includes a useful set of tools to perform standard, and more advanced, analyses of leaf gas exchange data.
 The open source framework combined with version control allows further development of the code.
@@
 # Acknowledgments {.unnumbered}
-David Ellsworth, Belinda Medlyn, Martin de Kauwe are acknowledged for discussion on fitting A-Ci curves.
-Special thanks to Belinda Medlyn whose code in the MAESTRA model provided inspiration for an early version of the Photosyn function.
+David Ellsworth, Belinda Medlyn and Martin de Kauwe are acknowledged for discussions on fitting A-C~i~ curves.
+Special thanks to Belinda Medlyn whose code in the MAESTRA model provided the basis for an early version of the Photosyn function.
 Thanks to John Drake for testing.

### 2015-09-08 R/functions.R
Added draft new figure (not in function yet), more plotting functions

@@ -28,3 +28,74 @@ plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
   
   text(x,y,txt,font=2,...)
 }
+
+addpoly <- function(x,y1,y2,col=alpha("lightgrey",0.8),...){
+  ii <- order(x)
+  y1 <- y1[ii]
+  y2 <- y2[ii]
+  x <- x[ii]
+  polygon(c(x,rev(x)), c(y1, rev(y2)), col=col, border=NA,...)
+}
+
+predline <- function(fit, from=NULL, to=NULL, poly=TRUE, ...){
+  
+  if(is.null(from))from <- min(fit$model[,2], na.rm=TRUE)
+  if(is.null(to))to <- max(fit$model[,2], na.rm=TRUE)
+  
+  newdat <- data.frame(X = seq(from,to, length=101))
+  
+  nm <- names(coef(fit))
+  names(newdat)[1] <- nm[length(nm)]
+  
+  if(poly){
+    pred <- as.data.frame(predict(fit, newdat, se.fit=TRUE, interval="confidence")$fit)
+    addpoly(newdat[[1]], pred$lwr, pred$upr)
+  }
+  ablinepiece(fit, from=from, to=to, ...)
+  
+}
+
+ablinepiece <- function(a=NULL,b=NULL,reg=NULL,from=NULL,to=NULL,...){
+  
+  # Borrowed from abline
+  if (!is.null(reg)) a <- reg
+  
+  if (!is.null(a) && is.list(a)) {
+    temp <- as.vector(coefficients(a))
+    from <- min(a$model[,2], na.rm=TRUE)
+    to <- max(a$model[,2], na.rm=TRUE)
+    
+    if (length(temp) == 1) {
+      a <- 0
+      b <- temp
+    }
+    else {
+      a <- temp[1]
+      b <- temp[2]
+    }
+  }
+  
+  segments(x0=from,x1=to,
+           y0=a+from*b,y1=a+to*b,...)
+  
+}
+
+alpha <- function (colour, alpha = NA) {
+  col <- col2rgb(colour, TRUE)/255
+  if (length(colour) != length(alpha)) {
+    if (length(colour) > 1 && length(alpha) > 1) {
+      stop("Only one of colour and alpha can be vectorised")
+    }
+    if (length(colour) > 1) {
+      alpha <- rep(alpha, length.out = length(colour))
+    }
[…]

### 2015-09-08 figures.R

@@ -120,3 +120,56 @@ figure4 <- function(){
   plotlabel("(b)","bottomleft")
 }
   
+
+
+
+
+tum <- read.csv("TumbarumbaGasex_ACis_Medlyn.csv")
+
+tumh <- subset(tum, PARi > 1400)
+
+
+# poor curves (no saturation)
+bad <- c("12","18","39","30","31")
+
+
+library(plantecophys)
+f <- fitacis(subset(tumh, !Curve %in% bad), "Curve")
+
+pdf("tmp.pdf")
+plot(f)
+dev.off()
+
+# lin <- read.csv("http://files.figshare.com/1886204/WUEdatabase_merged_Lin_et_al_2015_NCC.csv")
+# tumspot <- subset(lin, Datacontrib == "Belinda Medlyn")
+# saveRDS(tumspot, "tumspot.rds")
+
+tumspot <- readRDS("tumspot.rds")
+gfit <- fitBB(tumspot, varnames=list(ALEAF="Photo",GS="Cond",VPD="VPD",
+                                     Ca="CO2S"))
+tumspot$GSpred <- predict(gfit$fit, tumspot)
+g1 <- coef(gfit)[[2]]
+
+# (a)
+windows(8,8)
+par(mfrow=c(2,2))
+plot(f, "oneplot", addlegend=F, highlight="25", trans=F)
+
+with(coef(f), plot(Vcmax, Jmax,
+                   pch=19, 
+                   xlim=c(40,160),
+                   ylim=c(80,260)))
+predline(lm(Jmax ~ Vcmax, data=coef(f)))
+
+with(tumspot, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
+                   pch=19, col=alpha("black",0.3)))
+abline(0,1)
+
+with(tumspot,plot(VPD, Photo/Trmmol,
+                  pch=19, col=alpha("black",0.3)))
+curve(0.1*mean(tumspot$CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE)
+
+
+
+
+

### 2015-09-08 manuscript.Rmd

@@
 \
-Remko A. Duursma^1^
+Remko A. Duursma^1\*^
 
 \
-^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia
+^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Penrith, NSW, Australia
 
 \
-*Corresponding author*: 
+\* *Corresponding author*: 
 Remko Duursma
@@
 It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @medlyn2002].
-To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~) [@wullschleger1993].
+To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~, and optionally R~d~) [@wullschleger1993].
 
@@
 1a).
-A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i{\Gamma}^{∗})/(k_2 + C_i)$, where ${\Gamma}^{∗}$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
+A~c~ and A~j~ are non-linear functions of the chloroplastic CO~2~ concentration (C~c~), both of the form $k_1(C_c{\Gamma}^{∗})/(k_2 + C_c)$, where ${\Gamma}^{∗}$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~.
 The details of these functions and the temperature dependence of the various parameters are described elsewhere [@medlyn2002].
-We have here used C~i~ to denote [CO~2~] concentration in the leaf.
-Alternatively we may account for the diffusional limitation from intercellular airspaces to the chloroplast [@ethier2004], by calculating C~c~ as,
+
+In the practical application of the FVcB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange avoiding the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
+In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~ available.
+When g~m~ is known, C~c~ is calculated from C~i~ with Eq.
+2.
 
@@
 
-where g~m~ is the mesophyll conductance.
+where g~m~ is the mesophyll conductance (mol m^-2^ s^-1^).
+Although this method assumes that g~m~ is constant for a given leaf, it is well known that g~m~ responds dynamically to fluctuations in environmental drivers [@flexas2008] although some of the variation in g~m~ may due to artefacts related to (photo-)respiratory effects on measured g~m~ with standard methods [@tholen2012].
+Because no model has been developed to date that adequately captures the variation in g~m~, it is constant in the current implementation presented here.
+However, it can be used to study the effects of non-constant g~m~ on rates of photosynthesis and its response to environmental drivers.
 
@@
 (b) The intersection of the supply and demand curves of photosynthesis.
-The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
+The `Photosyn` function solves for C~c~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
 
 \
-Fitting the model to data requires some finesse because the model is a minimum function of two non-linear equations that is sometimes difficult to fit.
+Fitting the FVcB model to data requires some finesse because net photosynthesis is modelled as a minimum function of two non-linear equations that is sometimes difficult to fit.
 Moreover, sample sizes collected are often small due to time constraints in the field.
@@
 Through Eq.
-1, we have a dependency of photosynthesis on the availability of the substrate, C~i~.
-To estimate C~i~ itself, we need an estimate of the stomatal conductance to CO~2~.
+1, we have a dependency of photosynthesis on the availability of the substrate, C~c~.
+To estimate C~c~ itself, we need C~i~, which can be estimated when with stomatal conductance to CO~2~.
 From Fick's law, we can relate A~n~ to g~s~ and C~i~ as,
@@
 (3)
-$$ A_n = g_s(C_a - C_i) $$
+$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
 
 \
-where g~s~ is the conductance to CO~2~.
-We now have two equations for A~n~.
-The 'demand function' is given by Eq.
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
+We now have two equations for A~n~: the 'demand function' is given by Eq.
 1, and the 'supply function' in Eq.
 3.
-At steady state these two equation should be equal, which can be graphically shown as in Fig.
+At steady state these two equations should be equal, which can be graphically shown as in Fig.
 1b (cf.
@@
 \
-where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters, and f(D) a function that describes the response to the vapour pressure deficit (D, [@leuning1995]) or relative humidity [@ballberry1987].
+where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters, and f(D) can be one of many functions that describe the response to the vapour pressure deficit (D, [@leuning1995; @medlyn2011]) or relative humidity [@ballberry1987].
 An alternative approach to modelling g~s~ is through the hypothesis that stomata act optimally in the sense that they maximize photosynthesis while minimizing water loss.
@@
 1), `fitaci` fits the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity.
-The fit is achieved via standard non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R).
+The fit is achieved via standard non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R, see [@ritz2008]).
 The `fitaci` function includes methods to estimate appropriate starting values from the data, and attempts the fits along a wide range of possible starting values.
@@
 
+```{r eval=TRUE, echo=FALSE, message=FALSE}
+acidata1$PPFD <- 1800
+f <- suppressWarnings(fitaci(acidata1))
+```
+
 \
@@
 Colored lines indicate the two photosynthesis rates in the FvCB model.
+In the default mode, the `fitaci` function estimates Vcmax, Jmax and Rd from the fitted curve.
+Optionally, Rd is provided as an input, for example when it was measured separately.
+In this example, Vcmax is `r round(coef(f)[1],1)`, Jmax is `r round(coef(f)[2],1)` and Rd is `r round(coef(f)[3],1)` (all in units of $\mu$mol m^-2^ s^-1^).
 

### 2015-09-09 manuscript.Rmd
More revisions. Added start of example application.

@@
 *Short title*: The plantecophys R package
-
-<!---
-For Submission to PLOS one
-e.g. http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0061217
-http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003929
-http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003270
-and see: http://journals.plos.org/plosone/s/submission-guidelines
-(Methods, software, databases, and tools)
---->
 
@@
 
+# Download data needed for Figure 5.
+
 # Make PDFs
@@
 The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular [CO~2~], C~i~).
-It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @medlyn2002].
+It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @bernacchi2001; @medlyn2002].
 To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~, and optionally R~d~) [@wullschleger1993].
@@
 \
-We do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @medlyn2002]. But generally it is of the form,
+We do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @bernacchi2001; @medlyn2002]. But generally it is of the form,
 
@@
 
-In the practical application of the FVcB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange avoiding the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
-In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~ available.
+In the practical application of the FVcB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange data [@sharkey1982], and this avoids the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
+In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~.
 When g~m~ is known, C~c~ is calculated from C~i~ with Eq.
@@
 Although this method assumes that g~m~ is constant for a given leaf, it is well known that g~m~ responds dynamically to fluctuations in environmental drivers [@flexas2008] although some of the variation in g~m~ may due to artefacts related to (photo-)respiratory effects on measured g~m~ with standard methods [@tholen2012].
-Because no model has been developed to date that adequately captures the variation in g~m~, it is constant in the current implementation presented here.
-However, it can be used to study the effects of non-constant g~m~ on rates of photosynthesis and its response to environmental drivers.
+Because no model has been developed to date that adequately captures the variation in g~m~, it is a constant parameter in the implementation presented here.
+However, it can still be used to study the effects of non-constant g~m~ on rates of photosynthesis and its response to environmental drivers, as the parameter can be varied in model simulations.
 
@@
 where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
-We now have two equations for A~n~: the 'demand function' is given by Eq.
-1, and the 'supply function' in Eq.
-3.
+We now have two equations for A~n~: the 'demand function' (Eq.
+1), and the 'supply function' (Eq.
+3).
 At steady state these two equations should be equal, which can be graphically shown as in Fig.
@@
 4), but with a different D response function (f(D) in Eq.
-3) than the original Ball-Berry model.
+3) compared to the original Ball-Berry model.
 
@@
 
-The main tools included in the plantecophys package are to a) fit A-C~i~ curves, b) fit Ball-Berry type models, c) simulate from the coupled leaf gas exchange model and d) calculate the optimal stomatal conductance.
+The main tools included in the plantecophys package are to a) fit A-C~i~ curves to estimate V~cmax~, J~max~ and R~d~, b) fit Ball-Berry type models, c) simulate from the coupled leaf gas exchange model and d) calculate the optimal stomatal conductance.
 A list of the key functions in the package are summarized in Table 1.
@@
 
+All code used in this article (including the code to generate the actual article in markdown), can be downloaded (Zenodo link here).
+
 ## Results and Discussion
@@
 Instead of fitting the minimum function (Eq.
-1), `fitaci` fits the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity.
-The fit is achieved via standard non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R, see [@ritz2008]).
+1), `fitaci` fits the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity (Eq.
+5).
+
+\
+(5)
+$$A_m = \frac{A_c+A_j - \sqrt{(A_c+A_j)^2-4 \theta A_c A_j}}{2 \theta} - R_d$$
+
+where $\theta$ is a shape parameter, set to 0.9999, and A~m~ is the hyperbolic minimum of A~c~ and A~j~.
+
+The fit of the FvCB model to data is achieved with non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R, see [@ritz2008]).
 The `fitaci` function includes methods to estimate appropriate starting values from the data, and attempts the fits along a wide range of possible starting values.
@@
 \
+Required inputs are measurements of A~n~ and C~i~, and optionally leaf temperature (T~leaf~), and photosynthetically active radiation (PAR).
+Also required are estimates of Michaelis-Menten constants (K~c~, K~o~ or the combination K~m~) and $\Gamma^{*}$, if these are not available they are estimated from T~leaf~ with standard methods [@medlyn2002].
+
 In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) (and other parameters like $\Gamma^{*}$, K~c~ and K~o~) all depend non-linearly on T~leaf~.
 The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model (following [@medlyn2002]).
-Optionally, measured K~m~ and $\Gamma^{*}$ can be provided instead.
+Optionally, measured (or otherwise modelled) K~m~ and $\Gamma^{*}$ can be provided as input.
 
 \
-The function takes a dataframe as input, which includes measurements of A~n~, C~i~ and optionally leaf temperature (T~leaf~) and photosynthetically active radiation (PAR).
-It then returns a special object which has a print and plot method.
-It is easily used like this:
+The function takes a dataframe as input, which includes measurements of A~n~, C~i~ and optionally T~leaf~ and PAR, and is easily used like this:
 
@@
 \
-The output of the above example is shown in Fig. 2. Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.
+The output of the above example is shown in Fig.
+2.
+Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.
+I show this functionality in the example application further below.
 
@@
 Colored lines indicate the two photosynthesis rates in the FvCB model.
-In the default mode, the `fitaci` function estimates Vcmax, Jmax and Rd from the fitted curve.
-Optionally, Rd is provided as an input, for example when it was measured separately.
-In this example, Vcmax is `r round(coef(f)[1],1)`, Jmax is `r round(coef(f)[2],1)` and Rd is `r round(coef(f)[3],1)` (all in units of $\mu$mol m^-2^ s^-1^).
+In the default mode, the `fitaci` function estimates V~cmax~, J~max~ and R~d~ from the fitted curve.
+Optionally, R~d~ is provided as an input, for example when it was measured separately.
+In this example, V~cmax~ was estimated as `r round(coef(f)[1],1)`, J~max~ was `r round(coef(f)[2],1)` and R~d~ was `r round(coef(f)[3],1)` (all in units of $\mu$mol m^-2^ s^-1^).
 
@@
 This yields estimates of g~1~ and (optionally ) g~0~, which are necessary inputs to the coupled leaf gas exchange model.
-Note that, unlike the description in this paper, the user must provide stomatal conductance to H~2~O (not CO~2~) as input to the fitting process, which is the standard output of portable gas exchange instruments.
+Note that the user must provide stomatal conductance to H~2~O (not CO~2~) as input to the fitting process, which is the standard output of portable gas exchange instruments.
+This function is demonstrated in the example application further below.
 
@@
 The intersection of the supply and demand curves of photosynthesis (Fig.
-1b) gives the steady-state intercellular CO~2~ concentration.
+1b) gives the steady-state intercellular CO~2~ concentration (C~i~).
 This is solved by the `Photosyn` function.
@@
 In field conditions, D is always strongly positively related to T~leaf~.
-The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014].
+The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for simultaneously [@lin2012; @duursma2014].
 Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D (calculated following [@duursma2014]).
@@
 
-# Plot (produces Figure 3b minus the formatting)
+# Plot (produces Figure 3b minus the special formatting)
 with(run1, plot(Tleaf, ALEAF))
@@
 The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar hypothesis.
-It finds the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates corresponding A~n~ and g~s~.
+It finds the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates corresponding A~n~ and g~s~, thus avoiding the specification of an f(D) function as in Eq.
+4.
 This implementation was used by Medlyn et al.
@@
 [@buckley2014], who demonstrated that when a boundary layer is present, frequently an optimal g~s~ cannot be found.
+
+### An example application
+
+To demonstrate a practical application of the key functions in the package, I use data from Medlyn et al.
+[-@medlyn2007] on *Eucalyptus delegatensis*.
+Both A-C~i~ curves and 'spot gas exchange' data (i.e.
+leaf gas exchange measurements at prevailing environmental conditions) were collected.
+Using the `fitacis` function, it is straightforward to fit XXX curves to the measured A-C~i~ curves in one line of code, and make standard plots of the fitted curves (shown iin Fig.
+5a one option).
+The fitted coefficients are easily extracted, and used to plot a standard comparison of fitted Vcmax and Jmax values (Fig.
+5b), which show the typical correlation between the two.
+Next I fit the spot gas exchange data, yielding estimates of g1 and g0, and we can quickly compare modelled to measured g~s~ (Fig.
+5c), and compare the model prediction of instantaneous transpiration efficiency (A/E) to measurements along the variation in D (Fig.
+5d).
+This simple example application is available in the published repository (see Methods).
 

### 2015-09-09 figures.R

@@ -160,48 +160,44 @@ figure4 <- function(){
   
 
 
+figure5 <- function(){
+  
+  tum <- read.csv("TumbarumbaGasex_ACis_Medlyn.csv")
+  
+  tumh <- subset(tum, PARi > 1400)
+  
+  # poor curves (no saturation)
+  bad <- c("12","18","39","30","31")
+  
+  f <- fitacis(subset(tumh, !Curve %in% bad), "Curve")
+  
+  # lin <- read.csv("http://files.figshare.com/1886204/WUEdatabase_merged_Lin_et_al_2015_NCC.csv")
+  # tumspot <- subset(lin, Datacontrib == "Belinda Medlyn")
+  # saveRDS(tumspot, "tumspot.rds")
+  
+  tumspot <- readRDS("tumspot.rds")
+  gfit <- fitBB(tumspot, varnames=list(ALEAF="Photo",GS="Cond",VPD="VPD",
+                                       Ca="CO2S"))
+  tumspot$GSpred <- predict(gfit$fit, tumspot)
+  g1 <- coef(gfit)[[2]]
+  
+  # (a)
+  windows(8,8)
+  par(mfrow=c(2,2), mar=c(5,5,1,1), cex=1.1)
+  plot(f, "oneplot", addlegend=F, highlight="25", trans=F)
+  
+  with(coef(f), plot(Vcmax, Jmax,
+                     pch=19, 
+                     xlim=c(40,160),
+                     ylim=c(80,260)))
+  predline(lm(Jmax ~ Vcmax, data=coef(f)))
+  
+  with(tumspot, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
+                     pch=19, col=alpha("black",0.3)))
+  abline(0,1)
+  
+  with(tumspot,plot(VPD, Photo/Trmmol,
+                    pch=19, col=alpha("black",0.3)))
+  curve(0.1*mean(tumspot$CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE)
 
-tum <- read.csv("TumbarumbaGasex_ACis_Medlyn.csv")
-
-tumh <- subset(tum, PARi > 1400)
-
-
-# poor curves (no saturation)
-bad <- c("12","18","39","30","31")
-
-
-library(plantecophys)
-f <- fitacis(subset(tumh, !Curve %in% bad), "Curve")
-
-pdf("tmp.pdf")
-plot(f)
[…]

### 2015-09-09 manuscript.Rmd
Cleaned up organization; finished example application and new figure

@@
 \
-^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Penrith, NSW, Australia
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Penrith, NSW, Australia
 
@@
 Remko Duursma
-E: remkoduursma@gmail.com
-T: +61(0)422096908
+E: remkoduursma@westernsydney.edu.au
+T: +61(0)245701806
 
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
-
-# Tables in markdown
-library(ascii)
-
-# mine
-library(plantecophys)
-
-# use git
-library(git2r)
-
-# figures
-source("figures.R")
-
-# output
-if(!dir.exists("output"))dir.create("output")
-
-# Do calculations here, create objects that we can use throughout the paper.
-
-# Or, place them all in a separate script:
-# source("manuscript_calcs.R")
-
-# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
-f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))
-
-# Nice function to format P values.
-pval <- function(...)formatPval(..., includeEquality=TRUE)
-
-# Set Flag to include figures in text or not.
-includeFigs <- FALSE
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
+# Load packages and custom functions
+source("R/load.R")
 
@@
 
-# Download data needed for Figure 5.
+# Analysis for Figure 5.
+source("R/analysis.R")
 
 # Make PDFs
-to.pdf(figure1(), "output/figure1.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure2(), "output/figure2.pdf", width=4, height=4, verbose=FALSE)
-to.pdf(figure3(), "output/figure3.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure4(), "output/figure4.pdf", width=8, height=4, verbose=FALSE)
+source("R/make_figures.R")
+
+# This document can be compiled with the 'Knit Word' button in Rstudio, or:
+# library(rmarkdown)
+# render("manuscript.Rmd", "word_document", "manuscript.docx")
 ```
@@
 
+\
 In the practical application of the FVcB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange data [@sharkey1982], and this avoids the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
@@
 
+\
 where g~m~ is the mesophyll conductance (mol m^-2^ s^-1^).
@@
 
+\
 All code used in this article (including the code to generate the actual article in markdown), can be downloaded (Zenodo link here).
@@
 
+\
 where $\theta$ is a shape parameter, set to 0.9999, and A~m~ is the hyperbolic minimum of A~c~ and A~j~.
-
 The fit of the FvCB model to data is achieved with non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R, see [@ritz2008]).
@@
 
-To demonstrate a practical application of the key functions in the package, I use data from Medlyn et al.
+To demonstrate a practical application of the key functions in the package, I use field-collected data from Medlyn et al.
 [-@medlyn2007] on *Eucalyptus delegatensis*.
@@
 leaf gas exchange measurements at prevailing environmental conditions) were collected.
-Using the `fitacis` function, it is straightforward to fit XXX curves to the measured A-C~i~ curves in one line of code, and make standard plots of the fitted curves (shown iin Fig.
-5a one option).
-The fitted coefficients are easily extracted, and used to plot a standard comparison of fitted Vcmax and Jmax values (Fig.
+Using the `fitacis` function, it is straightforward to fit all `r length(acifits)` curves to the A-C~i~ data, and make standard plots of the fitted curves (shown in Fig.
+5a).
+The fitted coefficients are easily extracted, and used to plot a standard comparison of fitted V~cmax~ and J~max~ values (Fig.
 5b), which show the typical correlation between the two.
-Next I fit the spot gas exchange data, yielding estimates of g1 and g0, and we can quickly compare modelled to measured g~s~ (Fig.
-5c), and compare the model prediction of instantaneous transpiration efficiency (A/E) to measurements along the variation in D (Fig.
+Next I fit the spot gas exchange data, yielding estimates of g~1~ and (optionally) g~0~, and we can quickly compare modelled to measured g~s~ (Fig.
+5c), and compare the model prediction of instantaneous transpiration efficiency (A~n~/E) to measurements along the variation in D (Fig.
 5d).
 This simple example application is available in the published repository (see Methods).
+The simplified code for this example (panels a-c), omitting special formatting and a few minor settings is given below.
+
+```
+# Fit A-Ci curves. In this case, each separate curve is indexed by a column names 'Curve',
+# and the data were already read into a dataframe (tumh)
+acifits <- fitacis(tumh, "Curve")
+
+# Plot all A-Ci curves in one panel, highlight one fitted curve.
+plot(acifits, "oneplot", highlight="25")
+
+# Fit Medlyn et al.'s (2011) version of the Ball-Berry model
+# Data are already read into a dataframe (tumspot), and have standard names (or they can be set).
+gfit <- fitBB(tumspot, gsmodel="BBOpti")
+
+# Plot measured versus modelled, by predicting from the fitted model.
+tumspot$GSpred <- predict(gfit$fit, tumspot)
+with(tumspot, plot(GSpred, Cond))
+```
+
+\
+**Figure 5**.
+Example application of the plantecophys package to A-C~i~ curves and spot gas exchange measurements on *Eucalyptus delegatensis*.
+(a) Fitted A-C~i~ curves with one curve highlighted (b) Estimates of J~max~ plotted against V~cmax~, obtained from the fitted curves in panel (a).
+Solid line is a regression line (J~max~ = `r round(lmjvt$estimate[1],2)` + `r signif(lmjvt$estimate[2],2)` V~cmax~, R^2^ = `r signif(lmjvg$r.squared,2)`) with a 95% confidence interval for the mean.
+(c) Modelled (with the model of Medlyn et al.
+2011) versus measured g~s~ (p `r pval(lmgpredg$p.value)`, R^2^ = `r signif(lmgpredg$r.squared,2)`).
+Measurements included a wide range of environmental conditions (PAR, T~leaf~, D).
+In this example, only g~1~ was fit (estimate = `r signif(g1,3)`, 95% CI = `r paste(signif(g1ci,3), collapse=" - ")`).
+(d) The predicted response of ITE (A~n~ / E) as a function of D from the fitted model in panel (c) (solid line), and the measurements from panel (c) when PAR > 1000.
 

### 2015-09-09 R/analysis.R

@@ -1,0 +1,40 @@
+
+# Download data needed for Figure 5.
+tum <- read.csv("TumbarumbaGasex_ACis_Medlyn.csv")
+tumh <- subset(tum, PARi > 1400)
+
+# lin <- read.csv("http://files.figshare.com/1886204/WUEdatabase_merged_Lin_et_al_2015_NCC.csv")
+# tumspot <- subset(lin, Datacontrib == "Belinda Medlyn")
+# saveRDS(tumspot, "tumspot.rds")
+tumspot <- readRDS("tumspot.rds")
+
+
+# Fit A-Ci curves for Figure 5.
+# poor curves (no saturation)
+bad <- c("12","18","39")
+
+cf <- "cache/acifits.rds"
+if(!file.exists(cf)){
+  acifits <- fitacis(subset(tumh, !Curve %in% bad), "Curve")
+  saveRDS(acifits, cf)
+} else {
+  acifits <- readRDS(cf)
+}
+
+# Fit BB model to spot gas exchange data for Figure 5.
+gfit <- fitBB(tumspot, gsmodel="BBOpti",varnames=list(ALEAF="Photo",GS="Cond",VPD="VPD",
+                                                          Ca="CO2S"))
+tumspot$GSpred <- predict(gfit$fit, tumspot)
+g1 <- coef(gfit)[[2]]
+g1ci <- suppressMessages(unname(confint(gfit$fit)))
+
+lmgpred <- lm(Cond ~ GSpred, data=tumspot)
+lmgpredt <- tidy(lmgpred)
+lmgpredg <- glance(lmgpred)
+
+
+# Regression of Jmax vs. Vcmax (used in caption for Figure 5).
+p <- coef(acifits)
+lmjv <- lm(Jmax ~ Vcmax, data=p)
+lmjvt <- tidy(lmjv)
+lmjvg <- glance(lmjv)

### 2015-09-09 R/functions.R

@@ -1,3 +1,10 @@
+# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
+f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))
+
+# Nice function to format P values.
+pval <- function(...)formatPval(..., includeEquality=TRUE)
+
+
 to.pdf <- function(expr, filename, ..., verbose=TRUE) {
   if(!file.exists(dirname(filename)))
     dir.create(dirname(filename), recursive=TRUE)

### 2015-09-09 R/load.R

@@ -1,0 +1,22 @@
+# Two useful packages for printing numbers, accessing content of objects.
+library(broom)
+library(reporttools)
+
+# Tables in markdown
+library(ascii)
+
+# mine
+library(plantecophys)
+
+# use git
+library(git2r)
+
+# figures
+source("figures.R")
+
+# output
+if(!dir.exists("output"))dir.create("output")
+
+# cache, for downloaded files and analyses
+if(!dir.exists("cache"))dir.create("cache")
+

### 2015-09-09 R/make_figures.R

@@ -1,0 +1,5 @@
+to.pdf(figure1(), "output/figure1.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure2(), "output/figure2.pdf", width=4, height=4, verbose=FALSE)
+to.pdf(figure3(), "output/figure3.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure4(), "output/figure4.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure5(), "output/figure5.pdf", width=8, height=8, verbose=FALSE)

### 2015-09-09 figures.R

@@ -5,8 +5,8 @@ library(plantecophys)
 
 figure1 <- function(){
   
-  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1,
-      mfrow=c(1,2))
+  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(4,4,1,1), cex.axis=1,
+      mfrow=c(1,2), mgp=c(2,0.5,0), las=1, tcl=0.2)
   
   ci <- seq(40, 1050, length=101)
   acirun <- Aci(Ci=ci, Vcmax=65, Jmax=130, Rd=1.5)
@@ -33,7 +33,6 @@ figure1 <- function(){
   # PAnel b
   acirun <- Aci(Ci=ci)
   
-  par(xaxs="i", yaxs="i", cex.lab=1.3, mar=c(5,5,1,1), cex.axis=1)
   with(acirun, plot(Ci, ALEAF, type='l',
                     ylim=c(0,20), xlim=c(0,800),
                     lwd=2,
@@ -64,8 +63,9 @@ figure2 <- function(){
   acidata1$PPFD <- 1800
   f <- suppressWarnings(fitaci(acidata1))
   
-  par(mar=c(5,5,1,1), cex.lab=1.3, cex.axis=1, xaxs="i", yaxs="i")
-  plot(f, xlim=c(0,1500), ylim=c(0,30))
+  par(mar=c(4,4,1,1), cex.lab=1.3, cex.axis=1, xaxs="i", yaxs="i",
+      mgp=c(2,0.5,0), las=1, tcl=0.2)
+  plot(f, xlim=c(0,1500), ylim=c(0,30), legendbty='n')
   
 }
 
@@ -94,7 +94,8 @@ figure3 <- function(){
   p <- Photosyn(Tleaf=tairs, VPD=vpdfun(tairs))
   
   
-  par(mfrow=c(1,2), mar=c(5,5,1,1), xaxs="i", yaxs="i", cex.lab=1.2)
+  par(mfrow=c(1,2), mar=c(4,4,1,1), xaxs="i", yaxs="i", cex.lab=1.2,
+      mgp=c(2,0.5,0), las=1, tcl=0.2)
   plot(1, xlim=c(0,500), ylim=c(0,20), type='n',
        xlab=expression(italic(C)[i]~~(ppm)),
        ylab=expression(italic(A)[n]~~(mu*mol~m^-2~s^-1)))
[…]

### 2015-09-09 manuscript.Rmd
Edited language section

@@
 
-## Language
+## Language and availability
 
@@
 As such it builds easily, and is highly portable.
+A stable release is available on CRAN.
 The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys).
@@
 \
-All code used in this article (including the code to generate the actual article in markdown), can be downloaded (Zenodo link here).
+All code used in this article (including the code to generate the actual article in markdown), can be downloaded from http://www.github.com/duursmaplantecophys2015.
 

### 2015-09-09 manuscript.Rmd
Edits to response letter; minor editorial fixes to code and ms

--- 2015-09-09 manuscript.Rmd
@@
 \
-\* *Corresponding author*: 
-Remko Duursma
-E: remkoduursma@westernsydney.edu.au
-T: +61(0)245701806
+\* *Corresponding author*
+Email: r.duursma@westernsydney.edu.au
 
@@
 As such it builds easily, and is highly portable.
-The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys).
+The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys), from which a development version of the package can easily be installed.
 The repository includes an issue tracker, where users can suggest changes or report bugs.
@@
 \
-All code used in this article (including the code to generate the actual article in markdown), can be downloaded (Zenodo link here).
+All code used in this article (including the code to generate the article written in markdown, all figures and full example code), can be downloaded from http://www.github.com/duursma2015plosone.
 
-## Results and Discussion
+# Results and Discussion
 
@@
 \
+Because the fitting uses non-linear least squares, standard methods can be employed to estimate standard errors (SE), confidence intervals, and correlation of the fitted parameters.
+The fitaci function returns by default the SE and confidence intervals, and the built-in help page for the fitaci function shows how the nlstools package can be used to provide a detailed overview of the statistics of the non-linear least squares fit.
+
+\
 Required inputs are measurements of A~n~ and C~i~, and optionally leaf temperature (T~leaf~), and photosynthetically active radiation (PAR).
-Also required are estimates of Michaelis-Menten constants (K~c~, K~o~ or the combination K~m~) and $\Gamma^{*}$, if these are not available they are estimated from T~leaf~ with standard methods [@medlyn2002].
-
+Also required are estimates of Michaelis-Menten constants (K~c~, K~o~ or the combination K~m~) and $\Gamma^{*}$.
 In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) (and other parameters like $\Gamma^{*}$, K~c~ and K~o~) all depend non-linearly on T~leaf~.
@@
 acidata1$PPFD <- 1800
-f <- suppressWarnings(fitaci(acidata1))
+myfit <- suppressWarnings(fitaci(acidata1))
+fitc <- myfit$pars
 ```
@@
 Optionally, R~d~ is provided as an input, for example when it was measured separately.
-In this example, V~cmax~ was estimated as `r round(coef(f)[1],1)`, J~max~ was `r round(coef(f)[2],1)` and R~d~ was `r round(coef(f)[3],1)` (all in units of $\mu$mol m^-2^ s^-1^).
+In this example, V~cmax~ was estimated as `r round(fitc["Vcmax","Estimate"],1)` (SE `r round(fitc["Vcmax","Std.
+Error"],2)`), J~max~ was `r round(fitc["Jmax","Estimate"],1)` (SE `r round(fitc["Jmax","Std.
+Error"],2)`) and R~d~ was `r round(fitc["Rd","Estimate"],1)` (SE `r round(fitc["Rd","Std.
+Error"],2)`) (all in units of $\mu$mol m^-2^ s^-1^).
 
@@
 Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration.
-In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda E$ versus the C~i~, demonstrating that an optimum exists.
-The `FARAO` function finds this optimum numerically and calculates A~n~ and g~s~.
+In (a), individual curves at a range of values for the vapour pressure deficit (D) are plots of $A-\lambda E$ as a function of C~i~, demonstrating that an optimum C~i~ exists.
+The `FARAO` function finds this optimum numerically and calculates corresponding A~n~ and g~s~.
 The corresponding response of g~s~ to D is shown in panel (b).
@@
 # Acknowledgments {.unnumbered}
-David Ellsworth, Belinda Medlyn and Martin de Kauwe are acknowledged for discussions on fitting A-C~i~ curves.
-Special thanks to Belinda Medlyn whose code in the MAESTRA model provided the basis for an early version of the Photosyn function.
-Thanks to John Drake for testing.
+David Ellsworth, Belinda Medlyn and Martin De Kauwe are acknowledged for discussions on fitting A-C~i~ curves.
+Special thanks to Belinda Medlyn for the code in the MAESTRA model which provided the basis for an early version of the Photosyn function, and for sharing the data used in the example application.
+Thanks to John Drake and Nick Smith for testing and suggesting new features.
 

### 2015-09-09 R/make_figures.R

@@ -1,5 +1,5 @@
-to.pdf(figure1(), "output/figure1.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure2(), "output/figure2.pdf", width=4, height=4, verbose=FALSE)
-to.pdf(figure3(), "output/figure3.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure4(), "output/figure4.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure5(), "output/figure5.pdf", width=8, height=8, verbose=FALSE)
+to.pdf(figure1(), "output/Fig1.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure2(), "output/Fig2.pdf", width=4, height=4, verbose=FALSE)
+to.pdf(figure3(), "output/Fig3.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure4(), "output/Fig4.pdf", width=8, height=4, verbose=FALSE)
+to.pdf(figure5(), "output/Fig5.pdf", width=8, height=8, verbose=FALSE)

### 2015-09-10 manuscript.Rmd
Minor fixes

@@
 \* *Corresponding author*
+\
 Email: r.duursma@westernsydney.edu.au
 
+\
 *Short title*: The plantecophys R package
@@
 Here I present the R package 'plantecophys', a toolkit to analyse and model leaf gas exchange data.
-Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models.
+Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analysed with a few key models.
 These models include the Farquhar-von Caemmerer-Berry (FvCB) model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf.
@@
 \
-We do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @bernacchi2001; @medlyn2002]. But generally it is of the form,
+I do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @bernacchi2001; @medlyn2002]. But generally it is of the form,
 
@@
 \
-In the practical application of the FVcB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange data [@sharkey1982], and this avoids the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
+In the practical application of the FvCB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange data [@sharkey1982], and it avoids the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
 In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~.
@@
 \
-Fitting the FVcB model to data requires some finesse because net photosynthesis is modelled as a minimum function of two non-linear equations that is sometimes difficult to fit.
+Fitting the FvCB model to data requires some finesse because net photosynthesis is modelled as a minimum function of two non-linear equations that is sometimes difficult to fit.
 Moreover, sample sizes collected are often small due to time constraints in the field.
 A widely used published method requires the user to specify the transition of V~cmax~ to J~max~ limitation [@sharkey2007], a process that is both arbitrary and prevents batch analysis.
-Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process (following [@gu2010]).
+Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process (following [@gu2010]), and does not report standard errors of the estimated parameters.
 Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]).
@@
 \
-Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyze data and perform simulations are scattered and subject to little standardization.
+Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyse data and perform simulations are scattered and subject to little standardization.
 Fitting the FvCB model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult.
@@
 The straightforward `fitBB` function provides an interface to non-linear or linear regression to fit one of three stomatal conductance models [@ballberry1987; @leuning1995; @medlyn2011].
-This yields estimates of g~1~ and (optionally ) g~0~, which are necessary inputs to the coupled leaf gas exchange model.
+This yields estimates of g~1~ and (optionally) g~0~, which are necessary inputs to the coupled leaf gas exchange model.
 Note that the user must provide stomatal conductance to H~2~O (not CO~2~) as input to the fitting process, which is the standard output of portable gas exchange instruments.
@@
 Transpiration is calculated with the Penman-Monteith equation, which accounts for boundary layer effects.
-The details of `PhotosynEB` are not described here (we refer to the documentation), because it is very similar to other implementations [@wang1998; @buckley2014].
+The details of `PhotosynEB` are not described here (see the built in help file for more information), because it is very similar to other implementations [@wang1998; @buckley2014].
 
@@
 ```
-# Fit A-Ci curves. In this case, each separate curve is indexed by a column names 'Curve',
+# Fit A-Ci curves. 
+# In this case, each separate curve is indexed by a column named 'Curve',
 # and the data were already read into a dataframe (tumh)
@@
 # Fit Medlyn et al.'s (2011) version of the Ball-Berry model
-# Data are already read into a dataframe (tumspot), and have standard names (or they can be set).
+# Data are already read into a dataframe (tumspot), 
+# and have standard names (or they can be set).
 gfit <- fitBB(tumspot, gsmodel="BBOpti")
@@
 
-We need an open source set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
+We need an open source set of tools to analyse leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
 At the moment there are no publicly available tools to fit A-C~i~ curves that can be used as part of a reproducable workflow.

### 2015-09-10 R/analysis.R
Use Library to load packages; download proper data from figshare

@@ -1,12 +1,16 @@
 
-# Download data needed for Figure 5.
-tum <- read.csv("TumbarumbaGasex_ACis_Medlyn.csv")
+
+# Read data for Figure 5.
+# See : http://figshare.com/articles/Tumbarumba_Gas_Exchange/1538079
+
+# See load.R for code to download the data to the cache/ subdirectory.
+
+# A-Ci curve data
+tum <- read.csv(acifn)
 tumh <- subset(tum, PARi > 1400)
 
-# lin <- read.csv("http://files.figshare.com/1886204/WUEdatabase_merged_Lin_et_al_2015_NCC.csv")
-# tumspot <- subset(lin, Datacontrib == "Belinda Medlyn")
-# saveRDS(tumspot, "tumspot.rds")
-tumspot <- readRDS("tumspot.rds")
+# Spot gas exchange data
+tumspot <- read.csv(spotfn)
 
 
 # Fit A-Ci curves for Figure 5.
@@ -38,3 +42,4 @@ p <- coef(acifits)
 lmjv <- lm(Jmax ~ Vcmax, data=p)
 lmjvt <- tidy(lmjv)
 lmjvg <- glance(lmjv)
+

### 2015-09-10 R/load.R

@@ -1,15 +1,24 @@
+
+source("R/utils.R")
+
 # Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
+Library(broom)
+Library(reporttools)
 
 # Tables in markdown
-library(ascii)
+Library(ascii)
 
 # mine
-library(plantecophys)
+Library(plantecophys)
+if(packageVersion("plantecophys") < "0.6.5")
+  stop("Please install plantecophys 0.6.5 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
 
 # use git
-library(git2r)
+Library(git2r)
+
+# for downloading files
+Library(downloader)
+
 
 # figures
 source("figures.R")
@@ -20,3 +29,19 @@ if(!dir.exists("output"))dir.create("output")
 # cache, for downloaded files and analyses
 if(!dir.exists("cache"))dir.create("cache")
 
+# Download Medlyn data for example application
+acifn <- "cache/MedlynAci.csv"
+spotfn <- "cache/MedlynSpot.csv"
+
+if(!file.exists(acifn)){
+  download("http://files.figshare.com/2256196/TumbarumbaGasex_ACis_Medlyn.csv", 
+         acifn)
+}
+if(!file.exists(spotfn)){
+  download("http://files.figshare.com/2256198/TumbarumbaGasex_Spot_Medlyn.csv",
+         spotfn)
+}
+
+
+
+         

### 2015-09-10 R/utils.R

@@ -1,0 +1,13 @@
+Library <- function(pkg, ...){
+  
+  PACK <- .packages(all.available=TRUE)
+  pkgc <- deparse(substitute(pkg))
+  
+  if(pkgc %in% PACK){
+    library(pkgc, character.only=TRUE)
+  } else {
+    install.packages(pkgc, ...)
+  }
+  
+}
+

### 2015-09-10 manuscript.Rmd
Made code work with data downloaded from figshare; edits to example application

@@
 Because the fitting uses non-linear least squares, standard methods can be employed to estimate standard errors (SE), confidence intervals, and correlation of the fitted parameters.
-The fitaci function returns by default the SE and confidence intervals, and the built-in help page for the fitaci function shows how the nlstools package can be used to provide a detailed overview of the statistics of the non-linear least squares fit.
+The `fitaci` function returns by default the SE and confidence intervals, and the built-in help page for the `fitaci` function shows how the `nlstools` package can be used to provide a detailed overview of the statistics of the non-linear least squares fit.
 
@@
 5b), which show the typical correlation between the two.
-Next I fit the spot gas exchange data, yielding estimates of g~1~ and (optionally) g~0~, and we can quickly compare modelled to measured g~s~ (Fig.
+
+\
+Next, I fit Eq.
+4 to the spot gas exchange data, yielding an estimate of g~1~ (Fig.
+5c).
+It is straightforward to compare modelled to measured g~s~ (Fig.
 5c), and compare the model prediction of instantaneous transpiration efficiency (A~n~/E) to measurements along the variation in D (Fig.
 5d).
-This simple example application is available in the published repository (see Methods).
-The simplified code for this example (panels a-c), omitting special formatting and a few minor settings is given below.
+Because `fitBB` can fit a number of Ball-Berry type variants, the various models can be easily compared in terms of goodness of fit.
+This simple example application is available in the published repository (see Methods), and simplified code for this example (panels a-c only), omitting special formatting and a few minor settings, is given below.
 
@@
 David Ellsworth, Belinda Medlyn and Martin De Kauwe are acknowledged for discussions on fitting A-C~i~ curves.
-Special thanks to Belinda Medlyn for the code in the MAESTRA model which provided the basis for an early version of the Photosyn function, and for sharing the data used in the example application.
-Thanks to John Drake and Nick Smith for testing and suggesting new features.
+Special thanks to Belinda Medlyn for the code in the MAESTRA model which provided the basis for an early version of the `Photosyn` function, and for sharing the data used in the example application.
+Thanks to John Drake for testing and suggesting new features.
 

### 2015-09-10 R/analysis.R

@@ -26,8 +26,7 @@ if(!file.exists(cf)){
 }
 
 # Fit BB model to spot gas exchange data for Figure 5.
-gfit <- fitBB(tumspot, gsmodel="BBOpti",varnames=list(ALEAF="Photo",GS="Cond",VPD="VPD",
-                                                          Ca="CO2S"))
+gfit <- fitBB(tumspot, gsmodel="BBOpti")
 tumspot$GSpred <- predict(gfit$fit, tumspot)
 g1 <- coef(gfit)[[2]]
 g1ci <- suppressMessages(unname(confint(gfit$fit)))

### 2015-09-10 figures.R

@@ -14,7 +14,7 @@ figure1 <- function(){
   with(acirun, {
     plot(Ci, Ac-Rd, type='l',col="red",lwd=2, ylim=c(0,40),
          xlab=expression(italic(C)[i]~~(ppm)),
-         ylab=expression(italic(A)~~(mu*mol~m^-2~s^-1)),
+         ylab=expression(italic(A)[n]~~(mu*mol~m^-2~s^-1)),
                     xlim=c(0,1050))
     lines(Ci, Aj-Rd, col="blue", lwd=2)
     lines(Ci, ALEAF, lwd=2)
@@ -164,8 +164,6 @@ figure4 <- function(){
 
 figure5 <- function(){
   
-
-
   # (a)
   par(mfrow=c(2,2), mar=c(4,4,1,1), cex=1.1, xaxs="i", yaxs="i",las=1,
       tcl=0.2,mgp=c(2,0.5,0),
@@ -191,13 +189,15 @@ figure5 <- function(){
   abline(0,1)
   plotlabel("(c)", "topleft")
   
-  with(subset(tumspot,PARin > 1000), plot(VPD, Photo/Trmmol,
+  with(subset(tumspot,PARi > 1000), 
+       {plot(VpdL, Photo/Trmmol,
                     xlab="D (kPa)",
                     xlim=c(0,3), ylim=c(0,14),
                     ylab=expression(ITE~~(mu*mol~CO[2]/mmol~H[2]*O)),
-                    pch=19, col=alpha("black",0.3)))
-  curve(0.1*mean(tumspot$CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE,
-        from=min(tumspot$VPD), to=max(tumspot$VPD))
+                    pch=19, col=alpha("black",0.3))
+        curve(0.1*mean(CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE,
+          from=min(VpdL), to=max(VpdL))
+       })
   plotlabel("(d)", "topleft")
   
 }

### 2015-09-10 manuscript.Rmd
Added reference

@@
 To demonstrate a practical application of the key functions in the package, I use field-collected data from Medlyn et al.
-[-@medlyn2007] on *Eucalyptus delegatensis*.
+[-@medlyn2007, @medlynfigshare] on *Eucalyptus delegatensis*.
 Both A-C~i~ curves and 'spot gas exchange' data (i.e.
@@
 **Figure 5**.
-Example application of the plantecophys package to A-C~i~ curves and spot gas exchange measurements on *Eucalyptus delegatensis*.
-(a) Fitted A-C~i~ curves with one curve highlighted (b) Estimates of J~max~ plotted against V~cmax~, obtained from the fitted curves in panel (a).
+Example application of the plantecophys package to A-C~i~ curves and spot gas exchange measurements on *Eucalyptus delegatensis* (data can be downloaded from [@medlynfigshare]) (a) Fitted A-C~i~ curves with one curve highlighted (b) Estimates of J~max~ plotted against V~cmax~, obtained from the fitted curves in panel (a).
 Solid line is a regression line (J~max~ = `r round(lmjvt$estimate[1],2)` + `r signif(lmjvt$estimate[2],2)` V~cmax~, R^2^ = `r signif(lmjvg$r.squared,2)`) with a 95% confidence interval for the mean.

### 2015-10-06 manuscript.Rmd
tumspot split in two seasons and both are used; several reviewer comments incorporated into ms

@@
 (a) Lines are A-C~i~ curves simulated at a range of values for T~leaf~.
-Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014]).
+Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014] : D = 0.000605*T~air~^2.39^).
 Note that as T~leaf~ and D increase, C~i~ decreases.
@@
 As an alternative, the `PhotosynEB` function calculates T~leaf~ from the leaf energy balance.
-Transpiration is calculated with the Penman-Monteith equation, which accounts for boundary layer effects.
+Transpiration is calculated with the Penman-Monteith equation [@jones1992], which accounts for boundary layer effects.
 The details of `PhotosynEB` are not described here (see the built in help file for more information), because it is very similar to other implementations [@wang1998; @buckley2014].
@@
 5a).
-The fitted coefficients are easily extracted, and used to plot a standard comparison of fitted V~cmax~ and J~max~ values (Fig.
-5b), which show the typical correlation between the two.
+The fitted coefficients can be extracted using the `coef` function, and used to plot a comparison of fitted V~cmax~ and J~max~ values, which show the typical correlation between the two (Fig.
+5b).
 
@@
 5c).
-It is straightforward to compare modelled to measured g~s~ (Fig.
-5c), and compare the model prediction of instantaneous transpiration efficiency (A~n~/E) to measurements along the variation in D (Fig.
-5d).
+In this example, I used the model of Medlyn et al.
+[@medlyn2011], which is given by Eq.
+6 (in this example, I assumed g~0~ = 0)
+
+\
+(6)
+$$ g_s = g_0 + 1.6(1 + \frac{g_1}{\sqrt(D)})\frac{A_n}{C_a}$$
+
+In Fig.
+5c, modelled g~s~ is compared to measurements (Fig.
+5c).
+To compare the model prediction of instantaneous transpiration efficiency (A~n~/E) to measurements along the variation in D (Fig.
+5d), Eq.
+6 can be rearranged to give (cf.
+[@duursma2013], where it is assumed that g~0~ = 0)
+
+\
+(7)
+$$A_n/E = \frac{C_a P_a}{1.6 (g_1 D_s^k + D_s)}$$
+
 Because `fitBB` can fit a number of Ball-Berry type variants, the various models can be easily compared in terms of goodness of fit.
@@
 We need an open source set of tools to analyse leaf gas exchange data, as these data form a cornerstone of plant physiological ecology.
-At the moment there are no publicly available tools to fit A-C~i~ curves that can be used as part of a reproducable workflow.
+At the moment there are no publicly available tools to fit A-C~i~ curves or perform simulations with the coupled leaf gas exchange model that can be used as part of a reproducable workflow.
 The plantecophys R package is implemented in widely used language for data analysis.

### 2015-10-06 R/analysis.R

@@ -12,6 +12,13 @@ tumh <- subset(tum, PARi > 1400)
 # Spot gas exchange data
 tumspot <- read.csv(spotfn)
 
+tumspot$Date <- as.Date(tumspot$Date, format="%d/%m/%Y")
+tumspot$Season <- factor(format(tumspot$Date, "%m-%Y"))
+
+# Choose only one season (the one with the best fit, of course)
+# tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
+tumspot_1 <- subset(tumspot, Season == "11-2001")
+tumspot_2 <- subset(tumspot, Season == "05-2002")
 
 # Fit A-Ci curves for Figure 5.
 # poor curves (no saturation)
@@ -26,15 +33,25 @@ if(!file.exists(cf)){
 }
 
 # Fit BB model to spot gas exchange data for Figure 5.
-gfit <- fitBB(tumspot, gsmodel="BBOpti")
-tumspot$GSpred <- predict(gfit$fit, tumspot)
-g1 <- coef(gfit)[[2]]
-g1ci <- suppressMessages(unname(confint(gfit$fit)))
+gfit_1 <- fitBB(tumspot_1, gsmodel="BBOpti")
+gfit_2 <- fitBB(tumspot_2, gsmodel="BBOpti")
 
-lmgpred <- lm(Cond ~ GSpred, data=tumspot)
-lmgpredt <- tidy(lmgpred)
-lmgpredg <- glance(lmgpred)
+tumspot_1$GSpred <- predict(gfit_1$fit, tumspot_1)
+tumspot_2$GSpred <- predict(gfit_2$fit, tumspot_2)
 
+g1_1 <- coef(gfit_1)[[2]]
+g1ci_1 <- suppressMessages(unname(confint(gfit_1$fit)))
+
+g1_2 <- coef(gfit_2)[[2]]
+g1ci_2 <- suppressMessages(unname(confint(gfit_2$fit)))
+
+
+lmgpred_1 <- lm(Cond ~ GSpred, data=tumspot_1)
+lmgpred_2 <- lm(Cond ~ GSpred, data=tumspot_2)
+lmgpredt_1 <- tidy(lmgpred_1)
+lmgpredt_2 <- tidy(lmgpred_2)
+lmgpredg_1 <- glance(lmgpred_1)
+lmgpredg_2 <- glance(lmgpred_2)
 
 # Regression of Jmax vs. Vcmax (used in caption for Figure 5).
 p <- coef(acifits)

### 2015-10-06 figures.R

@@ -182,22 +182,30 @@ figure5 <- function(){
   predline(lm(Jmax ~ Vcmax, data=coef(acifits)))
   plotlabel("(b)", "topleft")
   
-  with(tumspot, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
+  with(tumspot_1, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
                      xlab=expression(Modelled~g[s]~~(mol~m^-2~s^-1)),
                      ylab=expression(Measured~g[s]~~(mol~m^-2~s^-1)),
                      pch=19, col=alpha("black",0.3)))
+  with(tumspot_2, points(GSpred, Cond, pch=19, col=alpha("red",0.3)))
   abline(0,1)
   plotlabel("(c)", "topleft")
   
-  with(subset(tumspot,PARi > 1000), 
-       {plot(VpdL, Photo/Trmmol,
+  with(subset(tumspot_1,PARi > 1000), {
+        plot(VpdL, Photo/Trmmol,
                     xlab="D (kPa)",
                     xlim=c(0,3), ylim=c(0,14),
                     ylab=expression(ITE~~(mu*mol~CO[2]/mmol~H[2]*O)),
                     pch=19, col=alpha("black",0.3))
-        curve(0.1*mean(CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE,
+        curve(0.1*mean(CO2S)/(1.6*(g1_1*sqrt(x) + x)), add=TRUE,
           from=min(VpdL), to=max(VpdL))
        })
+  with(subset(tumspot_2,PARi > 1000), {
+    points(VpdL, Photo/Trmmol,
+         pch=19, col=alpha("red",0.3))
+    curve(0.1*mean(CO2S)/(1.6*(g1_2*sqrt(x) + x)), add=TRUE,
+          from=min(VpdL), to=max(VpdL))
+  })
+  
   plotlabel("(d)", "topleft")
   
 }

### 2015-10-06 manuscript.Rmd
Expanded description of FARAO

@@
 
-The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar hypothesis.
-It finds the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates corresponding A~n~ and g~s~, thus avoiding the specification of an f(D) function as in Eq.
-4.
+The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar [@cowan1977] hypothesis that stomata respond to environmental drivers in order to maximize photosynthesis while minimizing water loss.
 This implementation was used by Medlyn et al.
 [@medlyn2011] to compare a simplified model of optimal stomatal conductance to the full numerical solution.
+
+To find optimal stomatal conductance, `FARAO` finds the C~i~ for which the quantity $A_n - \lambda E$ is maximal, where E is the leaf transpiration rate and $\lambda$ is the marginal cost of water (an empirical parameter related to g~1~, see [@medlyn2011]).
+A~n~ is calculated directly as a function of C~i~ via the FvCB model (Eq.
+1), g~s~ is calculated by rearranging Eq.
+3, and E is calculated assuming perfect coupling (thus $E = g_s D/P_a$, where P~a~ is atmospheric pressure).
+This numerical routine does not need specification of an f(D) function as in Eq.
+4, instead, this function is an emergent property.
 In Fig.
-4a, I have calculated $A - \lambda E$ across a range of C~i~ values, and for different values of D.
+4a, I have calculated $A_n - \lambda E$ across a range of C~i~ values, and for different values of D.
 The `FARAO` function calculates the optima of these curves, which can for example be used to study the response of stomatal conductance to D (Fig.

### 2015-10-06 figures.R
FARAO figure minor fix

@@ -144,7 +144,7 @@ figure4 <- function(){
        ylim=c(-1,10),
        xlim=c(0,400),
        xlab=expression(italic(C)[i]~~(ppm)),
-       ylab=expression(A-lambda*E~~(mu*mol~m^-2~s^-1)))
+       ylab=expression(A[n]-lambda*E~~(mu*mol~m^-2~s^-1)))
   for(i in 1:length(vpds))addcurve(vpds[i], pch=21, bg=Cols[i], cex=1.2)
   plotlabel("(a)","bottomleft")
   legend("topleft", c(expression(italic(D) == 0.5*kPa),expression(italic(D) == 3.5*kPa)),

### 2015-10-07 R/analysis.R
Once again only one curve for tumspot

@@ -17,8 +17,7 @@ tumspot$Season <- factor(format(tumspot$Date, "%m-%Y"))
 
 # Choose only one season (the one with the best fit, of course)
 # tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
-tumspot_1 <- subset(tumspot, Season == "11-2001")
-tumspot_2 <- subset(tumspot, Season == "05-2002")
+tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
 
 # Fit A-Ci curves for Figure 5.
 # poor curves (no saturation)
@@ -33,25 +32,14 @@ if(!file.exists(cf)){
 }
 
 # Fit BB model to spot gas exchange data for Figure 5.
-gfit_1 <- fitBB(tumspot_1, gsmodel="BBOpti")
-gfit_2 <- fitBB(tumspot_2, gsmodel="BBOpti")
+gfit <- fitBB(tumspot, gsmodel="BBOpti")
+tumspot$GSpred <- predict(gfit$fit, tumspot)
+g1 <- coef(gfit)[[2]]
+g1ci <- suppressMessages(unname(confint(gfit$fit)))
 
-tumspot_1$GSpred <- predict(gfit_1$fit, tumspot_1)
-tumspot_2$GSpred <- predict(gfit_2$fit, tumspot_2)
-
-g1_1 <- coef(gfit_1)[[2]]
-g1ci_1 <- suppressMessages(unname(confint(gfit_1$fit)))
-
-g1_2 <- coef(gfit_2)[[2]]
-g1ci_2 <- suppressMessages(unname(confint(gfit_2$fit)))
-
-
-lmgpred_1 <- lm(Cond ~ GSpred, data=tumspot_1)
-lmgpred_2 <- lm(Cond ~ GSpred, data=tumspot_2)
-lmgpredt_1 <- tidy(lmgpred_1)
-lmgpredt_2 <- tidy(lmgpred_2)
-lmgpredg_1 <- glance(lmgpred_1)
-lmgpredg_2 <- glance(lmgpred_2)
+lmgpred <- lm(Cond ~ GSpred, data=tumspot)
+lmgpredt <- tidy(lmgpred)
+lmgpredg <- glance(lmgpred)
 
 # Regression of Jmax vs. Vcmax (used in caption for Figure 5).
 p <- coef(acifits)

### 2015-10-07 figures.R

@@ -182,15 +182,14 @@ figure5 <- function(){
   predline(lm(Jmax ~ Vcmax, data=coef(acifits)))
   plotlabel("(b)", "topleft")
   
-  with(tumspot_1, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
+  with(tumspot, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
                      xlab=expression(Modelled~g[s]~~(mol~m^-2~s^-1)),
                      ylab=expression(Measured~g[s]~~(mol~m^-2~s^-1)),
                      pch=19, col=alpha("black",0.3)))
-  with(tumspot_2, points(GSpred, Cond, pch=19, col=alpha("red",0.3)))
   abline(0,1)
   plotlabel("(c)", "topleft")
   
-  with(subset(tumspot_1,PARi > 1000), {
+  with(subset(tumspot,PARi > 1000), {
         plot(VpdL, Photo/Trmmol,
                     xlab="D (kPa)",
                     xlim=c(0,3), ylim=c(0,14),
@@ -199,12 +198,6 @@ figure5 <- function(){
         curve(0.1*mean(CO2S)/(1.6*(g1_1*sqrt(x) + x)), add=TRUE,
           from=min(VpdL), to=max(VpdL))
        })
-  with(subset(tumspot_2,PARi > 1000), {
-    points(VpdL, Photo/Trmmol,
-         pch=19, col=alpha("red",0.3))
-    curve(0.1*mean(CO2S)/(1.6*(g1_2*sqrt(x) + x)), add=TRUE,
-          from=min(VpdL), to=max(VpdL))
-  })
   
   plotlabel("(d)", "topleft")
   

### 2015-10-09 manuscript.Rmd
Semi-working remake workflow

@@
 }
+# 
+# # Some summaries of stats for Figure 5.
+# gfit <- fitBB(tumspot, gsmodel="BBOpti")
+# tumspot$GSpred <- predict(gfit$fit, tumspot)
+# g1 <- coef(gfit)[[2]]
+# g1ci <- suppressMessages(unname(confint(gfit$fit)))
+# 
+# lmgpred <- lm(Cond ~ GSpred, data=tumspot)
+# lmgpredt <- tidy(lmgpred)
+# lmgpredg <- glance(lmgpred)
+# 
+# # Regression of Jmax vs. Vcmax (used in caption for Figure 5).
+# p <- coef(acifits)
+# lmjv <- lm(Jmax ~ Vcmax, data=p)
+# lmjvt <- tidy(lmjv)
+# lmjvg <- glance(lmjv)
 
-# Analysis for Figure 5.
-source("R/analysis.R")
+acifits <- readRDS("cache/acifits.rds")
 
-# Make PDFs
-source("R/make_figures.R")
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
 ```
@@
 
-\
-**Figure 5**.
-Example application of the plantecophys package to A-C~i~ curves and spot gas exchange measurements on *Eucalyptus delegatensis* (data can be downloaded from [@medlynfigshare]) (a) Fitted A-C~i~ curves with one curve highlighted (b) Estimates of J~max~ plotted against V~cmax~, obtained from the fitted curves in panel (a).
-Solid line is a regression line (J~max~ = `r round(lmjvt$estimate[1],2)` + `r signif(lmjvt$estimate[2],2)` V~cmax~, R^2^ = `r signif(lmjvg$r.squared,2)`) with a 95% confidence interval for the mean.
-(c) Modelled (with the model of Medlyn et al.
-2011) versus measured g~s~ (p `r pval(lmgpredg$p.value)`, R^2^ = `r signif(lmgpredg$r.squared,2)`).
-Measurements included a wide range of environmental conditions (PAR, T~leaf~, D).
-In this example, only g~1~ was fit (estimate = `r signif(g1,3)`, 95% CI = `r paste(signif(g1ci,3), collapse=" - ")`).
-(d) The predicted response of ITE (A~n~ / E) as a function of D from the fitted model in panel (c) (solid line), and the measurements from panel (c) when PAR > 1000.
-
 # Conclusions

### 2015-10-09 R/analysis.R

@@ -1,49 +1,0 @@
-
-
-# Read data for Figure 5.
-# See : http://figshare.com/articles/Tumbarumba_Gas_Exchange/1538079
-
-# See load.R for code to download the data to the cache/ subdirectory.
-
-# A-Ci curve data
-tum <- read.csv(acifn)
-tumh <- subset(tum, PARi > 1400)
-
-# Spot gas exchange data
-tumspot <- read.csv(spotfn)
-
-tumspot$Date <- as.Date(tumspot$Date, format="%d/%m/%Y")
-tumspot$Season <- factor(format(tumspot$Date, "%m-%Y"))
-
-# Choose only one season (the one with the best fit, of course)
-# tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
-tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
-
-# Fit A-Ci curves for Figure 5.
-# poor curves (no saturation)
-bad <- c("12","18","39")
-
-cf <- "cache/acifits.rds"
-if(!file.exists(cf)){
-  acifits <- fitacis(subset(tumh, !Curve %in% bad), "Curve")
-  saveRDS(acifits, cf)
-} else {
-  acifits <- readRDS(cf)
-}
-
-# Fit BB model to spot gas exchange data for Figure 5.
-gfit <- fitBB(tumspot, gsmodel="BBOpti")
-tumspot$GSpred <- predict(gfit$fit, tumspot)
-g1 <- coef(gfit)[[2]]
-g1ci <- suppressMessages(unname(confint(gfit$fit)))
-
-lmgpred <- lm(Cond ~ GSpred, data=tumspot)
-lmgpredt <- tidy(lmgpred)
-lmgpredg <- glance(lmgpred)
-
-# Regression of Jmax vs. Vcmax (used in caption for Figure 5).
-p <- coef(acifits)
-lmjv <- lm(Jmax ~ Vcmax, data=p)
-lmjvt <- tidy(lmjv)
-lmjvg <- glance(lmjv)
-

### 2015-10-09 R/data_processing.R

@@ -1,0 +1,46 @@
+
+# See : http://figshare.com/articles/Tumbarumba_Gas_Exchange/1538079
+process_acidata <- function(fn){
+
+  # A-Ci curve data
+  tum <- read.csv(acifn)
+  
+  # Use only light saturation (data include light response curves)
+  tumh <- subset(tum, PARi > 1400)
+  
+  # Fit A-Ci curves for Figure 5.
+  # poor curves (no saturation wrt Ci)
+  bad <- c("12","18","39")
+  tumh <- subset(tumh, !Curve %in% bad)
+  
+return(tumh)
+}
+
+process_spotdata <- function(fn){
+  # Spot gas exchange data
+  tumspot <- read.csv(spotfn)
+  
+  tumspot$Date <- as.Date(tumspot$Date, format="%d/%m/%Y")
+  tumspot$Season <- factor(format(tumspot$Date, "%m-%Y"))
+  
+  # Choose only one season (the one with the best fit, of course)
+  # tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
+  tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
+
+return(tumspot)
+}
+
+fit_medlynacicurves <- function(tumh){
+  
+  acifits <- fitacis(tumh, "Curve")
+return(acifits)
+}
+
+fit_medlynBB <- function(tumspot){
+
+  # Fit BB model to spot gas exchange data for Figure 5.
+  gfit <- fitBB(tumspot, gsmodel="BBOpti")
+
+return(gfit)
+}
+

### 2015-10-09 R/figures.R

@@ -1,7 +1,4 @@
 # Figures
-source("R/functions.R")
-library(plantecophys)
-
 
 figure1 <- function(){
   
@@ -83,9 +80,6 @@ figure3 <- function(){
     with(r1, lines(Ci, ALEAF, type='l', col=linecol, lwd=2))
     
     r2 <- Photosyn(Ca=400, VPD=vpd, Tleaf=tair)
-    #   gc <- r2$GS / 1.57
-    #   ablinepiece(gc*r2$Ca, -gc,  col="darkgrey",
-    #               from=r2$Ci, to=r2$Ca)
     points(r2$Ci, r2$ALEAF, pch=21, ...)
     
   }
@@ -103,10 +97,7 @@ figure3 <- function(){
                                   cex=1.2,  bg=Cols[i])
   box()
   plotlabel("(a)","topleft")
-#   for(i in 1:(nrow(p)-1)){
-#     arrows(x0=p$Ci[i], x1=p$Ci[i+1], y0=p$ALEAF[i], y1=p$ALEAF[i+1],
-#            col=Cols[i], length=0.15)
-#   }
+
   legend("left", c(expression(T[leaf] == 5),
                       expression(T[leaf] == 40)),
          lty=1, lwd=2, col=Cols[c(1,length(tairs))], bty='n', cex=0.8)
@@ -162,7 +153,7 @@ figure4 <- function(){
   
 
 
-figure5 <- function(){
+figure5 <- function(acifits, tumspot){
   
   # (a)
   par(mfrow=c(2,2), mar=c(4,4,1,1), cex=1.1, xaxs="i", yaxs="i",las=1,
@@ -173,6 +164,7 @@ figure5 <- function(){
   box()
   plotlabel("(a)", "topleft")
   
+  # (b)
   with(coef(acifits), plot(Vcmax, Jmax,
                            xlab=expression(V[cmax]~~(mu*mol~m^-2~s^-1)),
                            ylab=expression(J[max]~~(mu*mol~m^-2~s^-1)),
@@ -182,6 +174,7 @@ figure5 <- function(){
   predline(lm(Jmax ~ Vcmax, data=coef(acifits)))
   plotlabel("(b)", "topleft")
   
+  # (c)
   with(tumspot, plot(GSpred, Cond, xlim=c(0,0.6), ylim=c(0,0.6),
                      xlab=expression(Modelled~g[s]~~(mol~m^-2~s^-1)),
[…]

### 2015-10-09 R/load.R

@@ -1,47 +1,47 @@
-
-source("R/utils.R")
-
-# Two useful packages for printing numbers, accessing content of objects.
-Library(broom)
-Library(reporttools)
-
-# Tables in markdown
-Library(ascii)
-
-# mine
-Library(plantecophys)
-if(packageVersion("plantecophys") < "0.6.5")
-  stop("Please install plantecophys 0.6.5 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
-
-# use git
-Library(git2r)
-
-# for downloading files
-Library(downloader)
-
-
-# figures
-source("figures.R")
-
-# output
-if(!dir.exists("output"))dir.create("output")
-
-# cache, for downloaded files and analyses
-if(!dir.exists("cache"))dir.create("cache")
-
-# Download Medlyn data for example application
-acifn <- "cache/MedlynAci.csv"
-spotfn <- "cache/MedlynSpot.csv"
-
-if(!file.exists(acifn)){
-  download("http://files.figshare.com/2256196/TumbarumbaGasex_ACis_Medlyn.csv", 
-         acifn)
-}
-if(!file.exists(spotfn)){
-  download("http://files.figshare.com/2256198/TumbarumbaGasex_Spot_Medlyn.csv",
-         spotfn)
-}
-
-
-
-         
+# 
+# source("R/utils.R")
+# 
+# # Two useful packages for printing numbers, accessing content of objects.
+# Library(broom)
+# Library(reporttools)
+# 
+# # Tables in markdown
+# Library(ascii)
+# 
+# # mine
+# Library(plantecophys)
+# if(packageVersion("plantecophys") < "0.6.5")
+#   stop("Please install plantecophys 0.6.5 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
+# 
+# # use git
+# Library(git2r)
+# 
+# # for downloading files
+# Library(downloader)
+# 
+# 
+# # figures
+# source("R/figures.R")
+# 
+# # output
+# if(!dir.exists("output"))dir.create("output")
+# 
+# # cache, for downloaded files and analyses
+# if(!dir.exists("cache"))dir.create("cache")
+# 
+# # Download Medlyn data for example application
+# acifn <- "cache/MedlynAci.csv"
+# spotfn <- "cache/MedlynSpot.csv"
+# 
+# if(!file.exists(acifn)){
[…]

### 2015-10-09 R/make_figures.R

@@ -1,5 +1,5 @@
-to.pdf(figure1(), "output/Fig1.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure2(), "output/Fig2.pdf", width=4, height=4, verbose=FALSE)
-to.pdf(figure3(), "output/Fig3.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure4(), "output/Fig4.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure5(), "output/Fig5.pdf", width=8, height=8, verbose=FALSE)
+# to.pdf(figure1(), "output/Fig1.pdf", width=8, height=4, verbose=FALSE)
+# to.pdf(figure2(), "output/Fig2.pdf", width=4, height=4, verbose=FALSE)
+# to.pdf(figure3(), "output/Fig3.pdf", width=8, height=4, verbose=FALSE)
+# to.pdf(figure4(), "output/Fig4.pdf", width=8, height=4, verbose=FALSE)
+# to.pdf(figure5(), "output/Fig5.pdf", width=8, height=8, verbose=FALSE)

### 2015-10-09 R/make_ms.R

@@ -1,0 +1,3 @@
+make_ms <- function(fn){
+  rmarkdown::render(fn, word_document(), "manuscript.docx")
+}

### 2015-10-09 R/make_figures.R
minor bug fixes

@@ -2,4 +2,4 @@ to.pdf(figure1(), "output/Fig1.pdf", width=8, height=4, verbose=FALSE)
 to.pdf(figure2(), "output/Fig2.pdf", width=4, height=4, verbose=FALSE)
 to.pdf(figure3(), "output/Fig3.pdf", width=8, height=4, verbose=FALSE)
 to.pdf(figure4(), "output/Fig4.pdf", width=8, height=4, verbose=FALSE)
-to.pdf(figure5(), "output/Fig5.pdf", width=8, height=8, verbose=FALSE)
+to.pdf(figure5(acifits, tumspot), "output/Fig5.pdf", width=8, height=8, verbose=FALSE)

### 2015-10-09 figures.R

@@ -162,7 +162,7 @@ figure4 <- function(){
   
 
 
-figure5 <- function(){
+figure5 <- function(acifits, tumspot){
   
   # (a)
   par(mfrow=c(2,2), mar=c(4,4,1,1), cex=1.1, xaxs="i", yaxs="i",las=1,
@@ -195,7 +195,7 @@ figure5 <- function(){
                     xlim=c(0,3), ylim=c(0,14),
                     ylab=expression(ITE~~(mu*mol~CO[2]/mmol~H[2]*O)),
                     pch=19, col=alpha("black",0.3))
-        curve(0.1*mean(CO2S)/(1.6*(g1_1*sqrt(x) + x)), add=TRUE,
+        curve(0.1*mean(CO2S)/(1.6*(g1*sqrt(x) + x)), add=TRUE,
           from=min(VpdL), to=max(VpdL))
        })
   

### 2015-10-11 manuscript.Rmd
README update and minor cleanup

--- 2015-10-06 manuscript.Rmd
@@
 
-# Find current SHA of plantecophys
-rep_pkg <- "c:/repos/plantecophys"
-if(file.exists(rep_pkg)){
-  pkgrep <- repository(rep_pkg)
-  r <- reflog(pkgrep, verbose=FALSE)
-  cur_sha <- substr(r[[1]]@sha, 1, 7)
-} else {
-  cur_sha <- "???"
-}
+# Find current SHA of plantecophys (gives '???' on fail)
+cur_sha <- get_git_SHA("c:/repos/plantecophys")
 

### 2015-10-11 R/utils.R

@@ -11,3 +11,16 @@ Library <- function(pkg, ...){
   
 }
 
+
+get_git_SHA <- function(path){
+
+    if(file.exists(path)){
+    pkgrep <- repository(path)
+    r <- reflog(pkgrep, verbose=FALSE)
+    cur_sha <- substr(r[[1]]@sha, 1, 7)
+  } else {
+    cur_sha <- "???"
+  }
+
+return(cur_sha)
+}

### 2015-10-12 R/more_stats_acicurvefit.R
Added script with example code to extract more stats from fitted Aci curves

@@ -1,0 +1,50 @@
+
+#-- Additional statistics
+
+# The following bits of code show additional statistics that can be
+# extracted from fitted A-Ci curves.
+# Obviously, this list is not exhaustive.
+
+# I use this example curve fit:
+acidata1$PPFD <- 1800
+myfit <- fitaci(acidata1)
+
+# Show the usual output, including parameter estimates and confidence intervals.
+summary(myfit)
+
+# Correlation matrix of estimated Jmax, Vcmax and Rd from fitted A-Ci curve.
+# Example given for one curve fitted above.
+getCorMat_aci <- function(fit){
+  x <- summary(myfit$nlsfit, correlation=TRUE)
+  x$correlation
+}
+getCorMat_aci(myfit)
+
+
+# R2 of observed vs. fitted.
+getr2_acifit <- function(fit){
+  x <- lm(Ameas ~ Amodel, data=fit$df)
+  summary(x)$r.squared
+}
+
+getr2_acifit(myfit)
+
+# Residual plot (residual vs. fitted)
+residplot_acifit <- function(fit, ...){
+  
+  x <- myfit$df$Amodel
+  y <- with(myfit$df, Ameas - Amodel)
+  yl <- max(abs(y))
+  plot(x, y, ylim=c(-1.2*yl, 1.2*yl),
+       xlab="Fitted", ylab="Residual", ...)
+  abline(h=0)
+}
+residplot_acifit(myfit)
+
+
+
+
+
+
+
+

### 2015-10-12 manuscript.Rmd
2nd round revisions done, I think?

@@
 # render("manuscript.Rmd", "word_document", "manuscript.docx")
+# which assumes you have Pandoc installed (unless you run this from Rstudio),
+# and that you have an internet connection.
+
 ```
@@
 \
-In the practical application of the FvCB model, it is often assumed that C~c~ is equal to the intercellular [CO~2~] concentration (C~i~), which is useful because C~i~ can be estimated from leaf gas exchange data [@sharkey1982], and it avoids the need to estimate the conductance from intercellular airspaces to the chloroplast (mesophyll conductance, g~m~).
+In the practical application of the FvCB model, when leaf gas exchange is measured with a portable gas exchange instrument, estimates of C~c~ are difficult to obtain because they require an estimate of the mesophyll conductance.
+In this case, it is customary to us the intercellular [CO~2~] concentration (C~i~) as the driver of photosynthesis.
+This approach is useful because C~i~ can be estimated from concurrent measurements of CO~2~ and H~2~O flux [@sharkey1982].
 In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~.
-When g~m~ is known, C~c~ is calculated from C~i~ with Eq.
+When this is is known, C~c~ is calculated from C~i~ with Eq.
 2.
@@
 ```{r eval=TRUE, echo=FALSE, message=FALSE}
+# acidata1 is an example dataset in plantecophys (see ?acidata1)
 acidata1$PPFD <- 1800
-myfit <- suppressWarnings(fitaci(acidata1))
-fitc <- myfit$pars
+exfit1 <- suppressWarnings(fitaci(acidata1))
+fitc <- exfit1$pars
 ```
@@
 Error"],2)`) and R~d~ was `r round(fitc["Rd","Estimate"],1)` (SE `r round(fitc["Rd","Std.
-Error"],2)`) (all in units of $\mu$mol m^-2^ s^-1^).
+Error"],2)`).
+Assumed parameters were K~m~ = `r round(exfit1$Km,0)` and $\Gamma}^{∗}$ = `r round(exfit1$GammaStar,1)` (all in units of $\mu$mol m^-2^ s^-1^).
+The R^2^ of a regression of measured vs.
+fitted was 0.99.
 
@@
 Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D (calculated following [@duursma2014]).
-Figure 3b shows the output from `Photosyn`, which can be created with the following example.
-Note that in this example the default values of many parameters (e.g.
-J~max~, g~1~) are used in the call to `Photosyn`, but all of these can be set by the user.
+Both V~cmax~ and J~max~ have a peaked response to T~leaf~, so that at a given C~i~, A~n~ first increases with T~leaf and then decreases (lines, Fig.
+3a).
+As a result of increasing D, the modelled C~i~ decreases (symbols, Fig.
+3a).
+The net result is a peaked response of A~n~ as a function of D (Fig.
+3b).
 
@@
 Symbols are the same as in panel (a) but plotted against T~leaf~.
+
+The simplified code to produce Figure 3b, using the `Photosyn` function, is given below.
+Note that in this example the default values of many parameters (e.g.
+J~max~, g~1~) are used in the call to `Photosyn`, but all of these can be set by the user.
 

### 2015-10-12 R/analysis.R

@@ -5,22 +5,20 @@
 
 # See load.R for code to download the data to the cache/ subdirectory.
 
-# A-Ci curve data
+# A-Ci curve data (see load.R)
 tum <- read.csv(acifn)
 tumh <- subset(tum, PARi > 1400)
 
-# Spot gas exchange data
+# Spot gas exchange data (see load.R)
 tumspot <- read.csv(spotfn)
-
 tumspot$Date <- as.Date(tumspot$Date, format="%d/%m/%Y")
 tumspot$Season <- factor(format(tumspot$Date, "%m-%Y"))
 
-# Choose only one season (the one with the best fit, of course)
-# tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
+# Discard first season as it has a very different g1 (and g0).
 tumspot <- subset(tumspot, Season %in% c("11-2001","05-2002"))
 
 # Fit A-Ci curves for Figure 5.
-# poor curves (no saturation)
+# First remove these poor curves (no saturation)
 bad <- c("12","18","39")
 
 cf <- "cache/acifits.rds"
@@ -46,4 +44,10 @@ p <- coef(acifits)
 lmjv <- lm(Jmax ~ Vcmax, data=p)
 lmjvt <- tidy(lmjv)
 lmjvg <- glance(lmjv)
+
+
+
+
+
+
 

### 2015-10-12 R/load.R

@@ -10,8 +10,8 @@ Library(ascii)
 
 # mine
 Library(plantecophys)
-if(packageVersion("plantecophys") < "0.6.5")
-  stop("Please install plantecophys 0.6.5 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
+if(packageVersion("plantecophys") < "0.6.6")
+  stop("Please install plantecophys 0.6.6 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
 
 # use git
 Library(git2r)

### 2015-10-12 R/more_stats_acicurvefit.R

@@ -12,6 +12,12 @@ myfit <- fitaci(acidata1)
 # Show the usual output, including parameter estimates and confidence intervals.
 summary(myfit)
 
+# A summary of the underlying non-linear least-squares fit can be printed with,
+summary(myfit$nlsfit)
+
+# ... from which the RMSE can be easily extracted.
+summary(myfit$nlsfit)$sigma
+
 # Correlation matrix of estimated Jmax, Vcmax and Rd from fitted A-Ci curve.
 # Example given for one curve fitted above.
 getCorMat_aci <- function(fit){

### 2015-10-12 load.R
More logical organization

@@ -21,7 +21,8 @@ Library(downloader)
 
 
 # figures
-source("figures.R")
+# (Script with figure definitions, does not make PDFs)
+source("R/figures.R")
 
 # output
 if(!dir.exists("output"))dir.create("output")

### 2015-10-12 manuscript.Rmd
More logical organization
README update and minor edits

@@
 # Load packages and custom functions
-source("R/load.R")
+source("load.R")
 
@@
 This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991; @leuning1995] is implemented in many process-based ecosystem and global land surface models [@wang1998; @bonan2014; @duursma2012; @sellers1997].
-This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~, and their temperature dependencies).
+This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and incorporates key leaf traits (g~1~, V~cmax~, J~max~, R~d~, and their temperature dependencies).
 
@@
 The main tools included in the plantecophys package are to a) fit A-C~i~ curves to estimate V~cmax~, J~max~ and R~d~, b) fit Ball-Berry type models, c) simulate from the coupled leaf gas exchange model and d) calculate the optimal stomatal conductance.
-A list of the key functions in the package are summarized in Table 1.
+The key functions in the package are summarized in Table 1.
 
@@
 "Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
-"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvCB model when either C~i~ or g~s~ is given as input. For C4 photosynthesis, see `AciC4`",
+"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvCB model when either C~i~ or g~s~ is given as input.",
 "Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
@@
 All code used in this article (including the code to generate the article written in markdown, all figures and full example code), can be downloaded from http://www.github.com/duursma2015plosone.
+The repository also includes code to demonstrates how to extract additional statistics from fitted A-C~i~ curves.
 
@@
 Error"],2)`).
-Assumed parameters were K~m~ = `r round(exfit1$Km,0)` and $\Gamma}^{∗}$ = `r round(exfit1$GammaStar,1)` (all in units of $\mu$mol m^-2^ s^-1^).
+Assumed parameters were K~m~ = `r round(exfit1$Km,0)` and ${\Gamma}^{∗}$ = `r round(exfit1$GammaStar,1)` (all in units of $\mu$mol m^-2^ s^-1^).
 The R^2^ of a regression of measured vs.

### 2015-10-12 manuscript.Rmd
Final minor revisions

@@
 \
-In the practical application of the FvCB model, when leaf gas exchange is measured with a portable gas exchange instrument, estimates of C~c~ are difficult to obtain because they require an estimate of the mesophyll conductance.
+In the practical application of the FvCB model, when leaf gas exchange is measured with a portable gas exchange instrument, estimates of C~c~ are difficult to obtain because they require an estimate of the mesophyll conductance (g~m~).
 In this case, it is customary to us the intercellular [CO~2~] concentration (C~i~) as the driver of photosynthesis.
@@
 In the remainder of this article I will use C~c~ as the driver of photosynthesis, but point out that this can be replaced by C~i~ if the user does not have an estimate of g~m~.
-When this is is known, C~c~ is calculated from C~i~ with Eq.
+When g~m~ is known, C~c~ is calculated from C~i~ with Eq.
 2.
@@
 (b) The intersection of the supply and demand curves of photosynthesis.
-The `Photosyn` function solves for C~c~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
+The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.
 
@@
 This is solved by the `Photosyn` function.
-This flexible interface can be used to either 1) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`, equivalent to `Aci(...)`), 2) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf.
+This flexible interface can be used to either 1) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`; equivalent to `Aci(...)`), 2) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf.
 Fig.
@@
 Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D (calculated following [@duursma2014]).
-Both V~cmax~ and J~max~ have a peaked response to T~leaf~, so that at a given C~i~, A~n~ first increases with T~leaf and then decreases (lines, Fig.
+Both V~cmax~ and J~max~ have a peaked response to T~leaf~, so that at a given C~i~, A~n~ first increases with T~leaf~ and then decreases (lines, Fig.
 3a).
 As a result of increasing D, the modelled C~i~ decreases (symbols, Fig.
-3a).
+3a, as a consequence of Eq.
+4).
 The net result is a peaked response of A~n~ as a function of D (Fig.
@@
 
+\
 The simplified code to produce Figure 3b, using the `Photosyn` function, is given below.
@@
 
-To find optimal stomatal conductance, `FARAO` finds the C~i~ for which the quantity $A_n - \lambda E$ is maximal, where E is the leaf transpiration rate and $\lambda$ is the marginal cost of water (an empirical parameter related to g~1~, see [@medlyn2011]).
+To find optimal stomatal conductance, `FARAO` finds the C~i~ for which the quantity $A_n - \lambda E$ is maximal, where E is the leaf transpiration rate and $\lambda$ is the marginal cost of water (an empirical parameter related to g~1~, see [@cowan1977; @medlyn2011]).
 A~n~ is calculated directly as a function of C~i~ via the FvCB model (Eq.
@@
 To demonstrate a practical application of the key functions in the package, I use field-collected data from Medlyn et al.
-[-@medlyn2007, @medlynfigshare] on *Eucalyptus delegatensis*.
+[@medlyn2007; @medlynfigshare] on *Eucalyptus delegatensis*.
 Both A-C~i~ curves and 'spot gas exchange' data (i.e.

### 2017-09-12 R/utils.R
Updated README with reference etc.; reproducible again, thanks figshare and git2r for not caring

@@ -1,15 +1,3 @@
-Library <- function(pkg, ...){
-  
-  PACK <- .packages(all.available=TRUE)
-  pkgc <- deparse(substitute(pkg))
-  
-  if(pkgc %in% PACK){
-    library(pkgc, character.only=TRUE)
-  } else {
-    install.packages(pkgc, ...)
-  }
-  
-}
 
 
 get_git_SHA <- function(path){

### 2017-09-12 load.R

@@ -1,24 +1,14 @@
 
 source("R/utils.R")
 
-# Two useful packages for printing numbers, accessing content of objects.
-Library(broom)
-Library(reporttools)
 
-# Tables in markdown
-Library(ascii)
-
-# mine
-Library(plantecophys)
-if(packageVersion("plantecophys") < "0.6.6")
-  stop("Please install plantecophys 0.6.6 or newer, if needed from www.bitbucket.org/remkoduursma/plantecophys")
-
-# use git
-Library(git2r)
-
-# for downloading files
-Library(downloader)
-
+if(!require(pacman))install.packages("pacman")
+pacman::p_load(broom, reporttools,  # Two useful packages for 
+                                    #printing numbers, accessing content of objects.
+               ascii,                         #  Tables in markdown
+               plantecophys,
+               git2r,                         # use git from R
+               downloader)                    # for downloading files
 
 # figures
 # (Script with figure definitions, does not make PDFs)
@@ -31,17 +21,8 @@ if(!dir.exists("output"))dir.create("output")
 if(!dir.exists("cache"))dir.create("cache")
 
 # Download Medlyn data for example application
-acifn <- "cache/MedlynAci.csv"
-spotfn <- "cache/MedlynSpot.csv"
-
-if(!file.exists(acifn)){
-  download("http://files.figshare.com/2256196/TumbarumbaGasex_ACis_Medlyn.csv", 
-         acifn)
-}
-if(!file.exists(spotfn)){
-  download("http://files.figshare.com/2256198/TumbarumbaGasex_Spot_Medlyn.csv",
-         spotfn)
-}
+acifn <- "data/TumbarumbaGasex_ACis_Medlyn.csv"
+spotfn <- "data/TumbarumbaGasex_Spot_Medlyn.csv"
 
 
 
